---
title: "Geoexmap Processing"
author: "Malia Cortez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    code_folding: hide
---
## Setup
```{r setup, eval = FALSE}
# Process and save data to RDS file for use in geoexmap
# MC :)

# Processing:
## Filter raw datasets for just WA census tracts
## Use rasters to extract a mean or other aggregate for each census tract (rather than just using the raw rasters)
## Combine and put all measures into one very wide dataset, where each row is a census tract and contains all corresponding data
## Save to a RDS to conserve memory and use for geoexmap

# quickly loading data
library(data.table)
library(readxl)

# tidy and easy data manipulation
library(dplyr)
library(tidyverse)
library(lubridate)

# shiny
library(shiny)
library(bslib)

# LaTeX
#library(tinytex)

# raster/geospatial data manipulation and analysis
library(raster)
library(stars)
library(terra)
library(ncdf4)
library(sf)
library(sp)
library(tmaptools)
library(geosphere)
library(geojsonR)
library(spdep)
library(FNN)
library(exactextractr)

# TIGER shapefiles
library(tigris)

# querying Census data
library(tidycensus)

# calculating entropy-based segregation measures
library(segregation)
library(OasisR)

# plotting library for gut checking
library(ggplot2)
library(leaflet)
library(mapdeck)

# geocoding
library(tidygeocoder)

# web scraping
library(rvest)

# units
library(units)

# set working dir
#setwd("X:/fast/vopham_t/grp/geoexmap-geoexDC/geoexmap")
#library(tabulizer)
wa.tracts <- st_read("geoexmap/Geo/2020/wa_tracts_2020.gpkg")
```

## Geometry Processing

### Tract
```{r, eval=FALSE}
# load WA census tracts
wa.tracts.2010 <- tracts(state = "WA", year = 2010, cb = TRUE)
wa.tracts.2019 <- tracts(state = "WA", year = 2019, cb = TRUE)
wa.tracts.2020 <- tracts(state = "WA", year = 2020, cb = TRUE)
wa.tracts.2022 <- tracts(state = "WA", year = 2022, cb = TRUE)

wa.tracts.2015 <- tracts(state = "WA", year = 2015, cb = TRUE)

# same GEOIDs
setdiff(wa.tracts.2020$GEOID, wa.tracts.2019$GEOID)
setdiff(wa.tracts.2022$GEOID, wa.tracts.2020$GEOID)

plot(wa.tracts$geometry)

setdiff(wa.tracts.2019$GEOID, smoke$GEOID)
setdiff(smoke$GEOID, wa.tracts.2019$GEOID)

setdiff(wa.tracts$GEOID, smoke$GEOID)
setdiff(smoke$GEOID, wa.tracts$GEOID)

# extract multipolygons

crs(wa.tracts)

wa.tracts$HL_FIPS <- str_split_i(wa.tracts$NAME, "\\.", 1)
length(unique(wa.tracts$GEOID))
length(unique(wa.tracts$HL_FIPS))

length(unique(wa.tracts.2019$GEOID))
length(unique(smoke$GEOID))

wa.tracts.2019 <- st_transform(wa.tracts.2019, crs = 4326)
wa.tracts.2020 <- st_transform(wa.tracts.2020, crs = 4326)

st_write(wa.tracts.2019, "geoexmap/Geo/wa_tracts_2019.gpkg", append = FALSE)
st_write(wa.tracts.2020, "geoexmap/Geo/wa_tracts_2020.gpkg", append = FALSE)

```

```{r, eval=FALSE}
wa.tracts <- st_read("geoexmap/Geo/2020/wa_tracts_2020.gpkg")
wa.tracts.2019 <- st_read("geoexmap/Geo/2019/wa_tracts_2019.gpkg")
```

```{r, eval=FALSE}
plot(wa.tracts.2020$geometry)
```
```{r, eval=FALSE}
wa.tract.coords <- st_coordinates((st_centroid(st_geometry(wa.tracts))))
wa.tracts.sp <- as(wa.tracts, "Spatial")

neighbors <- poly2nb(wa.tracts.sp, row.names = wa.tracts$GEOID, queen = T)

# create weights/list
#list.queen <- nb2listw(neighbors, style = "W", zero.policy = TRUE)

#neighbors.queen <- listw2mat(list.queen)

neighbors.q <- nb2mat(neighbors, zero.policy = TRUE)
```

```{r, eval=FALSE}
plot(wa.tracts.sp)
plot(neighbors, wa.tract.coords, add = TRUE)
```

### City
```{r, eval=FALSE}
cities <- places(state = "WA", cb = TRUE)

plot(cities$geometry)

#st_write(cities, "geoexmap/Geo/city/cities.gpkg")
```


### County
```{r, eval=FALSE}
counties <- counties(state = "WA", cb = TRUE)

plot(counties$geometry)

st_write(counties, "geoexmap/Geo/county/counties.gpkg")
```

### Moran Test - Spatial Autocorrelation

```{r, eval=FALSE}
gmoran <- moran.test(caces.sp@data$no2, list.queen, alternative = "greater", na.action = na.exclude)
gmoran
gmoran[["p.value"]]
```
```{r, eval=FALSE}
gmoran <- moran.test(caces.sp@data$so2, list.queen, alternative = "greater", na.action = na.exclude)
gmoran
gmoran[["p.value"]]
```
```{r, eval=FALSE}
gmoran <- moran.test(caces.sp@data$co, list.queen, alternative = "greater", na.action = na.exclude)
gmoran
gmoran[["p.value"]]
```
All are significant, we reject the null in favor of the alternative hypothesis, supporting positive spatial autocorrelation.

```{r, eval=FALSE}
gmoranMC <- moran.mc(caces.sp@data$no2, list.queen, nsim = 999, na.action = na.exclude)
gmoranMC
hist(gmoranMC$res)
abline(v = gmoranMC$statistic, col = "red")
```
```{r, eval=FALSE}
gmoranMC <- moran.mc(caces.sp@data$so2, list.queen, nsim = 999, na.action = na.exclude)
gmoranMC
hist(gmoranMC$res)
abline(v = gmoranMC$statistic, col = "red")
```

```{r, eval=FALSE}
gmoranMC <- moran.mc(caces.sp@data$co, list.queen, nsim = 999, na.action = na.exclude)
gmoranMC
hist(gmoranMC$res)
abline(v = gmoranMC$statistic, col = "red")
```


## PLACES Outcomes, Behaviors, Prevention (2024, 2023 - Cervical Screening and Core Preventative Services for 65+)
```{r, eval=FALSE}
# ---- OUTCOMES ----
places.2024 <- fread("Data_Raw/PLACES__Census_Tract_Data__GIS_Friendly_Format___2024_release_20250324.csv") %>% 
  filter(StateDesc == "Washington")

places.2024$TractFIPS <- as.character(places.2024$TractFIPS)

places.2023 <- fread("Data_Raw/PLACES__Census_Tract_Data__GIS_Friendly_Format___2023_release_20250324.csv") %>% 
  filter(StateDesc == "Washington")

setdiff(places.2024$TractFIPS, wa.tracts.2020$GEOID)
setdiff(places.2023$TractFIPS, places.2024$TractFIPS)
setdiff(places.2024$TractFIPS, wa.tracts.2022$GEOID)

outcomes <- places.2024 %>% 
  dplyr::select(c(TractFIPS, TotalPopulation, ARTHRITIS_CrudePrev, CASTHMA_CrudePrev, BPHIGH_CrudePrev,
                  CANCER_CrudePrev, HIGHCHOL_CrudePrev, COPD_CrudePrev, CHD_CrudePrev, DEPRESSION_CrudePrev, 
                  DIABETES_CrudePrev, TEETHLOST_CrudePrev, STROKE_CrudePrev, OBESITY_CrudePrev))

behaviors <- places.2024 %>% 
  dplyr::select(c(TractFIPS, BINGE_CrudePrev, CSMOKING_CrudePrev, LPA_CrudePrev, SLEEP_CrudePrev))

prevention <- places.2024 %>% # missing cervical cancer screening, post-65 preventative care screening
  dplyr::select(c(TractFIPS, ACCESS2_CrudePrev, CHECKUP_CrudePrev, DENTAL_CrudePrev, BPMED_CrudePrev, CHOLSCREEN_CrudePrev,
                  MAMMOUSE_CrudePrev, COLON_SCREEN_CrudePrev))

social <- places.2024 %>% 
  dplyr::select(c(TractFIPS, FOODSTAMP_CrudePrev, FOODINSECU_CrudePrev, HOUSINSECU_CrudePrev, SHUTUTILITY_CrudePrev, LACKTRPT_CrudePrev,
                  EMOTIONSPT_CrudePrev))

outcomes$TractFIPS <- as.character(outcomes$TractFIPS)

outcomes.geo <- merge(outcomes, wa.tracts, by.x = "TractFIPS", by.y = "GEOID") 

behaviors.geo <- merge(behaviors, wa.tracts, by.x = "TractFIPS", by.y = "GEOID") %>% 
  dplyr::select(-c(UVI))

prevention.geo <- merge(prevention, wa.tracts, by.x = "TractFIPS", by.y = "GEOID")

social.geo <- merge(social, wa.tracts, by.x = "TractFIPS", by.y = "GEOID")

st_write(outcomes.geo, "geoexmap/Data_Processed/wa_outcomes.gpkg", append = FALSE)
st_write(behaviors.geo, "geoexmap/Data_Processed/wa_behaviors.gpkg")
st_write(prevention.geo, "geoexmap/Data_Processed/wa_prevention.gpkg")
st_write(social.geo, "geoexmap/Data_Processed/wa_social.gpkg")
```

## Outcomes: WSCR Incidence + Mortality Data
```{r, eval=FALSE}
# looks like if we skip some rows, we can get what we want
read_wscr <- function(path) {
  tryCatch({
    cat("\n Processing:", basename(path), "\n")
    
    # read lines to detect header/col names
    lines <- readLines(path, n = 30, warn = FALSE)
    
    keywords <- c("Data Type", "Cancer Site", "Geography", "Age Group", "Year", "Gender", "Race", "Observations")
    row.idx <- 0
    
    for (i in seq_along(lines)) {
      matches <- sum(str_detect(lines[i], regex(keywords, ignore_case = TRUE)))
      if (matches >= 3) {
        row.idx <- i
        cat("Header at line", i, "\n") # detect header
        break
      }
    }
    
    skip <- if (row.idx > 0) row.idx - 1 else 0
    cat("Skipping", skip, "lines\n")
    
    df <- read.csv(path, skip = skip)
    
    cat("File read! Dimensions:", nrow(df), "rows x", ncol(df), "columns\n")
    cat("Columns:", paste(colnames(df), collapse = ", "))
    
    return(df)
  }, error = function(e) {
    warning("Failed - error: ", e$message, "\n")
    return(NULL)
  }
  
  )
}
```

```{r, eval=FALSE}
wscr <- list.files("X:\\fast\\vopham_t\\grp\\geoexmap-geoexDC\\geoexmap\\geoexmap\\Data_Raw\\WSCR\\")
```

```{r, eval=FALSE}
setwd("X:\\fast\\vopham_t\\grp\\geoexmap-geoexDC\\geoexmap\\geoexmap\\Data_Raw\\WSCR\\")
all_data <- map(wscr, read_wscr)
```
```{r, eval=FALSE}
# data w 12 columns has stage at diagnosis
# data w 11 do not
# join to counties based on name
counties <- st_read("geoexmap/Geo/county/counties.gpkg")
```
```{r, eval=FALSE}
table(all$Geography)
```

```{r, eval=FALSE}
# some reported at multi county level
# map to individual counties
west <- c("Clallam", "Jefferson", "Grays Harbor", "Mason", "Thurston", "Pacific", "Lewis", "Cowlitz", "Skamania", "Clark")
kitsap_pierce <- c("Kitsap", "Pierce")
king <- c("King")
north <- c("Whatcom", "Skagit", "Snohomish", "Island", "San Juan")
central <- c("Chelan", "Douglas", "Grant", "Kittitas", "Yakima", "Benton", "Franklin", "Klickitat")
east <- c("Okanogan", "Ferry", "Stevens", "Pend Oreille", "Lincoln", "Spokane", "Adams", "Whitman", "Walla Walla", "Columbia", "Garfield", "Asotin")
```

```{r, eval=FALSE}
library(plyr)
incidence <- rbind(all_data[[1]], all_data[[2]], all_data[[3]])

all <- do.call(rbind.fill, all_data)
```

```{r, eval=FALSE}
# create county lookup list
county.lookup <- list(
  "West" = west,
  "Kitsap-Pierce" = kitsap_pierce,
  "King" = king,
  "North" = north,
  "Central" = central,
  "East" = east
)
```

```{r, eval=FALSE}
# expand the table, add counties for multi-county region
expanded <- all %>% 
  rowwise() %>% 
  mutate(
    counties = list(
      if (Geography %in% names(county.lookup)) {
        county.lookup[[Geography]]
      } else {
        Geography
      }
    )
  ) %>% 
  unnest(counties)
```

```{r, eval=FALSE}
wscr.exp <- expanded %>% 
  filter(Geography != "")

wscr.exp$Stage.At.Diagnosis[is.na(wscr.exp$Stage.At.Diagnosis)] <- "All"
```

```{r, eval=FALSE}
wscr.inc <- wscr.exp %>% 
  filter(Data.Type == "Incidence") %>% 
  dplyr::select(-Data.Type) #%>% 
  # pivot_wider(
  #   id_cols = counties,
  #   names_from = c(Cancer.Site, Stage.At.Diagnosis, Gender),
  #   values_from = Age.Adj..Rate.per.100.000,
  #   names_sep = "_"
  # )

wscr.mort <- wscr.exp %>% 
  filter(Data.Type == "Mortality") %>% 
  dplyr::select(-Data.Type)
```

```{r, eval=FALSE}
# join tables together in app?
#geo.wscr.inc <- merge(wscr.inc, counties, by.x = "COUNTY", by.y = "NAME", all.x = TRUE)
#geo.wscr.mort <- 
```
```{r, eval=FALSE}
fwrite(wscr.inc, "geoexmap/Data_Processed/wscr_inc.csv")
fwrite(wscr.mort, "geoexmap/Data_Processed/wscr_mort.csv")
```

## Sociodemographics

### ACS - Total Population, Age, Sex (Sam)
```{r, eval=FALSE}
tot.pop <- st_read("X:\\fast\\vopham_t\\grp\\geoexmap-geoexDC\\geoexmap\\geoexmap\\Data_Raw\\ACS_Age_Sex_Boundaries\\Tract.shp")

tot.pop <- tot.pop %>% 
  filter(State == "Washington")

tot.pop.tidy <- get_decennial(geography = "tract", table = "B01003", state = "Washington", geometry = TRUE)

#st_transform(crs = 4326)
```

### ACS - Race (Sam)
Race table shells:
- sex by age: B01001_001
- race: B02001_001
```{r, eval=FALSE}
# get_decennial
```

## Natural Environment
### PM 2.5
```{r natural-env, eval=FALSE}
#### PM 2.5 ####
pm2.5_2022_nc <- ncdfnc_open('Data_Raw/V5NA04.02.HybridPM25.NorthAmerica.2022-2022-JUL.nc')

{
  sink('V5NA04.02.HybridPM25.NorthAmerica.2022-2022-JUL.txt')
  print(pm2.5_2022_nc)
  sink()
}

# capture data in lat/long form
lon <- ncvar_get(pm2.5_2022_nc, "lon")
lat <- ncvar_get(pm2.5_2022_nc, "lat", verbose = FALSE)

# capture pm 2.5 data in an array
# measured in microgram/cubic meter
pm2.5_array_22 <- ncvar_get(pm2.5_2022_nc, "GWRPM25")

# capture data in lat/long form
lon <- ncvar_get(pm2.5_2020_nc, "lon")
lat <- ncvar_get(pm2.5_2020_nc, "lat", verbose = FALSE)

# capture pm 2.5 data in an array
# measured in microgram/cubic meter
pm2.5_array_20 <- ncvar_get(pm2.5_2020_nc, "GWRPM25")

# 2d array
dim(pm2.5_array_22)
dim(pm2.5_array_20)

nc_close(pm2.5_2022_nc)
nc_close(pm2.5_2020_nc)

# save to an array
pm2.5_raster_22 <- raster(t(pm2.5_array_22), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))

pm2.5_raster_20 <- raster(t(pm2.5_array_20), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))

# convert to spatraster
pm2.5_raster_22 <- as(pm2.5_raster_22, "SpatRaster")
pm2.5_raster_20 <- as(pm2.5_raster_20, "SpatRaster")

#plot(pm2.5_raster)

r <- flip(pm2.5_raster_22, direction = "vertical")
r2 <- flip(pm2.5_raster_20, direction = "vertical")

plot(r)

# check that projections are the same
st_crs(wa.tracts.2020) == st_crs(pm2.5_raster_22)
st_crs(wa.tracts.2020) == st_crs(pm2.5_raster_20)

# crop to WA extent
wa_pm2.5_22 <- crop(r, extent(wa.tracts.2020))
wa_pm2.5_20 <- crop(r2, extent(wa.tracts.2020))

par(mar=c(1, 1, 1, 1))
plot(wa_pm2.5_22)
plot(wa_pm2.5_20)

# extract mean value of the raster for each census tract
wa.tracts.2020$PM2.5 <- extract(wa_pm2.5_22, wa.tracts.2020, fun="mean", method = "simple", touches = FALSE,
                           na.rm = TRUE) 
wa.tracts.2020 <- wa.tracts.2020 %>% 
  #dplyr::select(c(GEOID, NAME, geometry, PM2.5$layer)) %>% 
  mutate(pm2.5_2022 = PM2.5$layer) %>% 
  dplyr::select(c(GEOID, NAME, pm2.5_2022))

wa.tracts.2020$PM2.5_2020 <- extract(wa_pm2.5_20, wa.tracts.2020, fun="mean", method = "simple", touches = FALSE,
                           na.rm = TRUE)
## DON'T NEED 2020
wa.tracts <- wa.tracts.2020 %>% 
  mutate(pm2.5_2020 = PM2.5_2020$layer) %>% 
  dplyr::select(c(GEOID, NAME, pm2.5_2022, pm2.5_2020))

st_write(wa.tracts, "geoexmap/Data_Processed/pm25_20-22.gpkg")
```

### Ozone (O3)
```{r, eval=FALSE}
ozone <- fread("Data_Raw/Daily_Census_Tract-Level_Ozone_Concentrations__2016_-_2020_20250324.csv")
```


### CACES: Nitrogen Dioxide (NO2), Sulfur Dioxide (SO2), Carbon Monoxide (CO) 
```{r, eval=FALSE}
# join 2019 tracts with caces
caces <- fread("Data_Raw/WA_LUR_2020_O3_CO_SO2_NO2_PM25.csv") %>% 
  pivot_wider(names_from = pollutant, values_from = pred_wght) %>% 
    dplyr::select(c(fips, no2, so2, co))

caces$fips <- as.character(caces$fips)

wa.2019 <- st_read("geoexmap/Geo/2019/wa_tracts_2019.gpkg")

caces.geo <- left_join(wa.2019, caces, by = c("GEOID" = "fips"))
```

```{r, eval=FALSE}
caces.coords <- st_coordinates((st_centroid(st_geometry(caces.geo))))
caces.sp <- as(caces.geo, "Spatial")

# add neighbors for disjointed polygons
addnbs <- function(sp.dat) {
  neighbors <- spdep::poly2nb(sp.dat, row.names = sp.dat$GEOID, queen = TRUE, snap = 500)

  count <- spdep::card(neighbors)

  if(!any(count == 0)) return(neighbors)
  
  # get nearest neighbor index, use centroids
  nnbs = spdep::knearneigh(coordinates(sp.dat))$nn
  
  no_edges_from <- which(count==0)
  
  for (i in no_edges_from) {
    neighbors[[i]] = nnbs[i]
  }
  
  return(neighbors)
}

neighbors <- addnbs(caces.sp)

# neighbors object to matrix
neighbors.q <- spdep::nb2mat(neighbors, zero.policy = TRUE)
```


```{r, eval=FALSE}
# still 8 disconnected polygons
spdep::n.comp.nb(neighbors)$nc
```


```{r, eval=FALSE}
spat_impute <- function(data, weights, var, coords, k = 5) {
  pollutant <- data[[var]]
  missing_idx <- which(is.na(pollutant))
  
  if (length(missing_idx) == 0) return(pollutant)
  
  # first pass
  # matrix where each row is the variable values repeated
  value_matrix <- matrix(rep(pollutant, each = length(missing_idx)), 
                        nrow = length(missing_idx))
  # weights for missing obs
  missing_weights <- weights[missing_idx, , drop = FALSE]
  missing_weights[is.na(value_matrix)] <- 0
  # row sums of weights
  weight_sums <- rowSums(missing_weights)
  # weighted sums
  weighted_sums <- rowSums(missing_weights * value_matrix, na.rm = TRUE)
  # weighted averages
  imputed_values <- ifelse(weight_sums > 0, weighted_sums / weight_sums, NA)
  
  # do a second pass: knn for remaining NA
  missing_idx_2 <- which(is.na(imputed_values))

  if (length(missing_idx_2) > 0) {
    library(FNN)

    complete <- which(!is.na(pollutant))

    if (length(complete) >= k) { # k: how many nearest neighbors to use when spatial neighbors missing
      # get coords for obs that are still missing
      missing.global <- missing_idx[missing_idx_2]
      missing.coords <- coords[missing.global, , drop = FALSE]

      knn.result <- get.knnx(coords[complete,],
                             missing.coords,
                             k = k)

      nearest.idx.mat <- matrix(complete[knn.result$nn.index],
                            nrow = length(missing_idx_2))
      nearest.dist.mat <- knn.result$nn.dist # get nearest neighbors with IDW

      # calc weights
      weights_matrix_knn <- 1 / (nearest.dist.mat + 1e-10) # prevent division by 0
      weights_sums_knn <- rowSums(weights_matrix_knn)
      weights_matrix_knn <- weights_matrix_knn / weights_sums_knn

      # neighbors for all missing obs
      neighbor_values_matrix <- matrix(pollutant[nearest.idx.mat],
                                       nrow = length(missing_idx_2))

      # calc weighted avg
      knn.imp <- rowSums(weights_matrix_knn * neighbor_values_matrix)

      imputed_values[missing_idx_2] <- knn.imp
    }
  }
  
  result <- pollutant
  result[missing_idx] <- imputed_values
  return(result)
}
```

```{r, eval=FALSE}
caces.geo$no2_imp <- spat_impute(caces.geo, neighbors.q, "no2", caces.coords)
caces.geo$so2_imp <- spat_impute(caces.geo, neighbors.q, "so2", caces.coords)
caces.geo$co_imp <- spat_impute(caces.geo, neighbors.q, "co", caces.coords)
```


```{r, eval=FALSE}
# gut check
og.missing <- is.na(caces.geo$no2)
caces.geo$imp <- ifelse(og.missing, "imputed", "observed")

ggplot(caces.geo) +
  geom_sf(aes(fill = no2_imp, color = imp), size = 0.3) +
  scale_fill_viridis_c(name = "Air Pollution") +
  scale_color_manual(values = c("imputed" = "red", "observed" = "transparent"), 
                     name = "Data Type") +
  theme_minimal() +
  labs(title = "Air Pollution: Observed vs Imputed Values")
```

```{r, eval=FALSE}
# transform to 2020 tracts (1772)
# steps from nhgis
## obtain data of interest for older geog
## join the crosswalk to older geog
## multiply older counts/values by crosswalk's interpolation weights, producing estimated counts for all intersections between older and newer geogs
## sum these atom counts for each new census tract (group by the 2020)

# get xwalk
xwalk <- fread("X:/fast/vopham_t/grp/geoexmap-geoexDC/geoexmap/geoexmap/nhgis_tr2010_tr2020_53.csv")
xwalk$tr2010ge <- as.character(xwalk$tr2010ge)

## obtain data of interest for older geog (caces)
# convert to sf
caces.wa <- caces.geo %>% 
  st_as_sf() %>% 
  dplyr::select(c(GEOID, no2_imp, so2_imp, co_imp))


## join the crosswalk to older geog
caces.wa.xwalk <- merge(xwalk, caces.wa, by.x = "tr2010ge", by.y = "GEOID", all.x = TRUE)

# using block groups for walkability
#caces.wa.xwalk <- merge(xwalk, walkability, by.x = "bg2010ge", by.y = "GEOID", all.x = TRUE)


## multiply older counts/values by crosswalk's interpolation weights, producing estimated counts for all intersections between older and newer geogs
caces.wa.xwalk$no2_2020 <- caces.wa.xwalk$parea * caces.wa.xwalk$no2_imp
caces.wa.xwalk$so2_2020 <- caces.wa.xwalk$parea * caces.wa.xwalk$so2_imp
caces.wa.xwalk$co_2020 <- caces.wa.xwalk$parea * caces.wa.xwalk$co_imp

## sum these atom counts for each new census tract (group by the 2020 tract fips)
# there's literally only 1 missing why !
length(unique(xwalk$tr2020ge))
setdiff(wa.tracts$GEOID, xwalk$tr2020ge)

caces.2020.wa <- caces.wa.xwalk %>% 
  group_by(tr2020ge) %>% 
  mutate(no2 = sum(no2_2020),
         so2 = sum(so2_2020),
         co = sum(co_2020)) %>% 
  slice(1) %>% 
  ungroup()

# manually replace: tr2020ge: 53033990100
caces.2020.wa$no2[caces.2020.wa$tr2020ge == 53033990100] <- 0.0
caces.2020.wa$so2[caces.2020.wa$tr2020ge == 53033990100] <- 0.0
caces.2020.wa$co[caces.2020.wa$tr2020ge == 53033990100] <- 0.0

caces.2020.wa <- caces.2020.wa %>% 
  filter(!is.na(no2)) %>% 
  dplyr::select(c(tr2020ge, no2, so2, co))

caces.wa.tracts <- merge(wa.tracts, caces.2020.wa, by.x = "GEOID", by.y = "tr2020ge", all.x = TRUE)
```

```{r, eval=FALSE}
# i guess we're using spatial weights again ??
caces.coords <- st_coordinates((st_centroid(st_geometry(caces.wa.tracts))))
caces.sp <- as(caces.wa.tracts, "Spatial")

neighbors <- addnbs(caces.sp)

# neighbors object to matrix
neighbors.q <- spdep::nb2mat(neighbors, zero.policy = TRUE)

caces.wa.tracts$no2_imp <- spat_impute(caces.wa.tracts, neighbors.q, "no2", caces.coords)
caces.wa.tracts$so2_imp <- spat_impute(caces.wa.tracts, neighbors.q, "so2", caces.coords)
caces.wa.tracts$co_imp <- spat_impute(caces.wa.tracts, neighbors.q, "co", caces.coords)

# yay!
caces.final <- caces.wa.tracts %>% 
  dplyr::select(c(GEOID, no2_imp, so2_imp, co_imp)) %>% 
  st_as_sf()

# finally!
st_write(caces.final, "geoexmap/Data_Processed/wa_no2_so2_co.gpkg")
```


### Wildfire Smoke / Near Surface Smoke (2023)
```{r, eval=FALSE}
# smoke from Stanford ECHO

# smoke.curr <- readRDS("../Archive/geoexmap_demo/Demo Data/smokePM2pt5_predictions_daily_tract_20060101-20231231.rds")
# wa.curr <- smoke.curr %>% 
#   filter(startsWith(GEOID, "53"))
# 
# smoke2020 <- wa.curr %>% 
#   filter(lubridate::year(date) == 2020)
# 
# smoke2021 <- wa.curr %>% 
#   filter(lubridate::year(date) == 2021)
# 
# fwrite(smoke2020, "Demo Data/wa_smoke_2020.csv")
# fwrite(smoke2021, "Demo Data/wa_smoke_2021.csv")

smoke23 <- fread("geoexmap/Data_Processed/wa_smoke_2023.csv")

smoke20$GEOID <- as.character(smoke20$GEOID)
smoke20$date <- as.Date(smoke20$date)

smoke21$GEOID <- as.character(smoke21$GEOID)
smoke21$date <- as.Date(smoke21$date)

smoke22$GEOID <- as.character(smoke22$GEOID)
smoke22$date <- as.Date(smoke22$date)

smoke23$GEOID <- as.character(smoke23$GEOID)
smoke23$date <- as.Date(smoke23$date)

setdiff(wa.tracts$GEOID, smoke20$GEOID)

length(unique(smoke23$GEOID))

wa.tracts$GEOID <- as.character(wa.tracts$GEOID)

wa.tracts.smoke <- wa.tracts %>% 
  dplyr::select(c("GEOID"))

# processing instructions from: https://www.dropbox.com/scl/fo/91k0aq80vp57qixkm508q/AIRVst386qB2G_N1XhnE4qk/tract?dl=0&preview=README.txt&rlkey=nutebc9pn2vsupr0p9ks4k73u&subfolder_nav_tracking=1
process_smoke <- function(preds) {
  dates <- seq.Date(ymd("20200101"), ymd("20231231"), by = "day")
  
  out <- expand.grid(GEOID = wa.tracts.smoke$GEOID, date = dates) # get full combination of tract days
  out <- left_join(out, preds, by = c("GEOID", "date")) # match smokePM predictions on smoke days to tract days
  
  out <- mutate(out, smokePM_pred = replace_na(smokePM_pred, 0))
  
  return(out)
}

smoke20.full <- process_smoke(smoke20)
smoke21.full <- process_smoke(smoke21)
smoke22.full <- process_smoke(smoke22)
smoke23.full <- process_smoke(smoke23)

# groupby geoid and calculate annual average
smoke_ann_avg <- function(dat) {
  return(dat %>% 
           group_by(GEOID) %>% 
           summarize(wildfirePM2.5 = mean(smokePM_pred)))
}

avg.wildfire2020 <- smoke_ann_avg(smoke20.full)
avg.wildfire2021 <- smoke_ann_avg(smoke21.full)
avg.wildfire2022 <- smoke_ann_avg(smoke22.full)
avg.wildfire2023 <- smoke_ann_avg(smoke23.full)

# join with tract data
wa.tracts.2020 <- merge(wa.tracts, avg.wildfire2020, by = "GEOID")
wa.tracts.2021 <- merge(wa.tracts.2020, avg.wildfire2021, by = "GEOID")
wa.tracts.2022 <- merge(wa.tracts.2021, avg.wildfire2022, by = "GEOID")
wa.tracts.2023 <- merge(wa.tracts, avg.wildfire2023, by = "GEOID")

# rename columns
wa.tracts <- wa.tracts.2023

names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.x"] <- "wildfirePM25_2020"
names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.y"] <- "wildfirePM25_2021"
names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.x.1"] <- "wildfirePM25_2022"
names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.y.1"] <- "wildfirePM25_2023"

st_write(wa.tracts, "geoexmap/Data_Processed/wildfire_smoke_2023.gpkg", append = FALSE)
```

### UV Radiation (2024)
```{r, eval=FALSE}
uv.world <- nc_open('Data_Raw/uvief2024_world.nc')

{
  sink('uvief2024_world.txt')
  print(uv.world)
  sink()
}

# capture data in lat/long form
lon <- ncvar_get(uv.world, "PRODUCT/longitude")
lat <- ncvar_get(uv.world, "PRODUCT/latitude", verbose = FALSE)

# capture pm 2.5 data in an array
# measured in microgram/cubic meter
uvi.array <- ncvar_get(uv.world, "PRODUCT/uvi_clear")

print(uvi.array)

# 3d array - latitude (1440) x longitude (720) x days (366 in 2024)
dim(uvi.array)

# take mean of 3rd dimension
mean.uvi <- rowMeans(uvi.array, dims = 2)

dim(mean.uvi)

nc_close(uv.world)

# save to an array
uvi.raster <- raster(t(mean.uvi), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))

uvi.spat <- as(uvi.raster, "SpatRaster")

#plot(pm2.5_raster)

r <- flip(uvi.spat, direction = "vertical")

plot(r)

# check that projections are the same
st_crs(wa.tracts.2020) == st_crs(uvi.spat)

# crop
wa.uvi <- crop(r, extent(wa.tracts.2020))

par(mar=c(1, 1, 1, 1))
plot(wa.uvi)

# extract mean value of the raster for each census tract
wa.tracts.2020$UVI <- extract(wa.uvi, wa.tracts.2020, fun="mean", method = "simple", touches = FALSE,
                           na.rm = TRUE) 

wa.tracts.2020 <- wa.tracts.2020 %>% 
  #dplyr::select(c(GEOID, NAME, geometry, PM2.5$layer)) %>% 
  mutate(UVI = UVI$layer) %>% 
  dplyr::select(c(GEOID, NAME, UVI))

leaflet(wa.tracts.2020) %>% 
  addTiles() %>% 
  addPolygons(
    stroke = TRUE,
    weight = 1,
    color = "black",
    fillOpacity = 0.7,
    smoothFactor = 0.5,
    fillColor = ~colorBin("YlOrRd", UVI)(UVI)
  )

st_write(wa.tracts.2020, "geoexmap/Data_Processed/wa_uvi.gpkg")
```


### National Risk Index
```{r, eval=FALSE}

```

### PRISM: Temperature, Precipitation, Dew Point
```{r, eval=FALSE}
 
```

### Radon 
```{r, eval=FALSE}
radon <- fread("Data_Raw/ZIPCODE_Average_Radon_2001_2021.csv") #%>% 
  #filter(STATEFP == 53)

zcta.tract <- fread("zcta_tract_rel_10.txt", sep = ",") #%>% 
  #filter(STATE == 53)

# join data to zcta tract
radon.xwalk <- merge(zcta.tract, radon, by.x = "ZCTA5", by.y = "ZIPCODE", all.x = TRUE)

radon.xwalk$screen_tr <- (radon.xwalk$TRHUPCT / 100) * radon.xwalk$screen_conc 

radon.tr <- radon.xwalk %>% 
  group_by(GEOID) %>% 
  mutate(screen_conc_tr = sum(screen_tr, na.rm = TRUE)) %>% 
  slice(1)

radon.nec <- radon.tr %>% 
  dplyr::select(c(GEOID, screen_conc_tr))

radon.nec$GEOID <- as.character(radon.nec$GEOID)

# join with wa data
# old geographies
#wa.radon <- merge(wa.tracts.2019, radon.nec, by.x = "GEOID", by.y = "GEOID", all.x = TRUE)
radon.geo <- left_join(wa.tracts.2019, radon.nec, by = c("GEOID" = "GEOID"))

summary(radon.geo)
```

```{r, eval=FALSE}
radon.coords <- st_coordinates((st_centroid(st_geometry(radon.geo))))
radon.sp <- as(radon.geo, "Spatial")

# add neighbors for disjointed polygons
addnbs <- function(sp.dat) {
  neighbors <- spdep::poly2nb(sp.dat, row.names = sp.dat$GEOID, queen = TRUE, snap = 500)

  count <- spdep::card(neighbors)

  if(!any(count == 0)) return(neighbors)
  
  # get nearest neighbor index, use centroids
  nnbs = spdep::knearneigh(coordinates(sp.dat))$nn
  
  no_edges_from <- which(count==0)
  
  for (i in no_edges_from) {
    neighbors[[i]] = nnbs[i]
  }
  
  return(neighbors)
}

neighbors <- addnbs(radon.sp)

# neighbors object to matrix
neighbors.q <- spdep::nb2mat(neighbors, zero.policy = TRUE)

spdep::n.comp.nb(neighbors)$nc
```
```{r, eval=FALSE}
spat_impute <- function(data, weights, var, coords, k = 5) {
  pollutant <- data[[var]]
  missing_idx <- which(is.na(pollutant))
  
  if (length(missing_idx) == 0) return(pollutant)
  
  # first pass
  # matrix where each row is the variable values repeated
  value_matrix <- matrix(rep(pollutant, each = length(missing_idx)), 
                        nrow = length(missing_idx))
  # weights for missing obs
  missing_weights <- weights[missing_idx, , drop = FALSE]
  missing_weights[is.na(value_matrix)] <- 0
  # row sums of weights
  weight_sums <- rowSums(missing_weights)
  # weighted sums
  weighted_sums <- rowSums(missing_weights * value_matrix, na.rm = TRUE)
  # weighted averages
  imputed_values <- ifelse(weight_sums > 0, weighted_sums / weight_sums, NA)
  
  # do a second pass: knn for remaining NA
  missing_idx_2 <- which(is.na(imputed_values))

  if (length(missing_idx_2) > 0) {
    library(FNN)

    complete <- which(!is.na(pollutant))

    if (length(complete) >= k) { # k: how many nearest neighbors to use when spatial neighbors missing
      # get coords for obs that are still missing
      missing.global <- missing_idx[missing_idx_2]
      missing.coords <- coords[missing.global, , drop = FALSE]

      knn.result <- get.knnx(coords[complete,],
                             missing.coords,
                             k = k)

      nearest.idx.mat <- matrix(complete[knn.result$nn.index],
                            nrow = length(missing_idx_2))
      nearest.dist.mat <- knn.result$nn.dist # get nearest neighbors with IDW

      # calc weights
      weights_matrix_knn <- 1 / (nearest.dist.mat + 1e-10) # prevent division by 0
      weights_sums_knn <- rowSums(weights_matrix_knn)
      weights_matrix_knn <- weights_matrix_knn / weights_sums_knn

      # neighbors for all missing obs
      neighbor_values_matrix <- matrix(pollutant[nearest.idx.mat],
                                       nrow = length(missing_idx_2))

      # calc weighted avg
      knn.imp <- rowSums(weights_matrix_knn * neighbor_values_matrix)

      imputed_values[missing_idx_2] <- knn.imp
    }
  }
  
  result <- pollutant
  result[missing_idx] <- imputed_values
  return(result)
}
```

```{r, eval=FALSE}
# add new variable with imputed values
radon.geo$rad_imp <- spat_impute(radon.geo, neighbors.q, "screen_conc_tr", radon.coords)
```

```{r, eval=FALSE}
xwalk <- fread("X:/fast/vopham_t/grp/geoexmap-geoexDC/geoexmap/geoexmap/nhgis_tr2010_tr2020_53.csv")
xwalk$tr2010ge <- as.character(xwalk$tr2010ge)

## obtain data of interest for older geog
# convert to sf
radon <- radon.geo %>% 
  dplyr::select(c(GEOID, rad_imp))


## join the crosswalk to older geog
radon.wa.xwalk <- merge(xwalk, radon, by.x = "tr2010ge", by.y = "GEOID", all.x = TRUE)
```

```{r, eval=FALSE}
## multiply older counts/values by crosswalk's interpolation weights, producing estimated counts for all intersections between older and newer geogs
radon.wa.xwalk$radon_2020 <- radon.wa.xwalk$wt_hu * radon.wa.xwalk$rad_imp
```

```{r, eval=FALSE}
radon.2020.wa <- radon.wa.xwalk %>% 
  #st_as_sf() %>% 
  group_by(tr2020ge) %>% 
  mutate(radon = sum(radon_2020, na.rm = TRUE)) %>% 
  slice(1) %>% 
  ungroup() %>%
  dplyr::select(c(tr2020ge, radon)) %>% 
  filter(startsWith(as.character(tr2020ge), "53")) 
  

radon.geo <- merge(wa.tracts, radon.2020.wa, by.x = "GEOID", by.y = "tr2020ge", all.x = TRUE) %>% 
  dplyr::select(c(GEOID, radon))
```

```{r, eval=FALSE}
# finally!
st_write(radon.geo, "geoexmap/Data_Processed/wa_radon.gpkg", append = FALSE)
```


### PFAS in Drinking Water
Steps:

- link contaminant result data with PWS boundaries
- then code with 0 = census tract doesn't overlap with water system with PFAS detected
```{r, eval=FALSE}
ucmr <- read_xlsx("Data_Raw/UCMR_Res_WA.xlsx")
pws <- fread("Data_Raw/PWS/Tracts.csv")
```

```{r, eval=FALSE}
wa.pws <- pws %>% 
  filter(startsWith(PWSID, "WA"))
```

```{r, eval=FALSE}
wa.ucmr <- merge(ucmr, wa.pws, by.x = "PWS ID", by.y = "PWSID")
```

```{r, eval=FALSE}
# filter out lithium, <MRL
wa.ucmr <- wa.ucmr %>% 
  filter(Contaminant != "lithium")
```

```{r, eval=FALSE}
# only detectable levels
wa.ucmr.num <- wa.ucmr[wa.ucmr[, 4] != "< MRL",]

wa.ucmr.num[, 4] <- as.numeric(wa.ucmr.num[,4])
```

```{r, eval=FALSE}
# reduce to 1 row per GEOID
# use mean, we won't be using values anyway

# rename so we can get column name
wa.ucmr.num <- wa.ucmr.num %>% 
  rename("Result" = 4)

wa.ucmr.tract <- wa.ucmr.num %>% 
  group_by(GEOID20) %>% 
  summarize(mean_result = mean(Result, na.rm = TRUE))

wa.tracts.ucmr <- merge(wa.tracts, wa.ucmr.tract, by.x = "GEOID", by.y = "GEOID20", all.x = TRUE)
```

```{r, eval=FALSE}
wa.tracts.ucmr$PFAS <- !is.na(wa.tracts.ucmr$mean_result)
```
`
```{r, eval=FALSE}
wa.tracts.ucmr <- wa.tracts.ucmr %>% 
  dplyr::select(c(GEOID, PFAS))
```

```{r, eval=FALSE}
st_write(wa.tracts.ucmr, "geoexmap/Data_Processed/wa_dw_pfas.gpkg", append = FALSE)
```

### Marine Microplastics
Retrieved from bounding box, find the largest distance to WA boundaries and use as threshold
```{r, eval=FALSE}
microplastics.wa <- fread("Data_Raw/Marine Microplastic Concentrations (2).csv")
micro.geo <- sf::st_as_sf(microplastics.wa, coords = c("Longitude(degree)", "Latitude (degree)")) %>%
  st_set_crs(crs(wa.tracts)) %>% 
  st_transform(crs = 4326)
```

```{r, eval=FALSE}
distances <- st_distance(micro.geo, wa.tracts)
```

```{r, eval=FALSE}
min.distances <- apply(distances, 1, min)
```

```{r, eval=FALSE}
max(min.distances)
```
 Included points less than 70398.2 meters away from Washington borders (bounding box).
 
```{r, eval=FALSE}
st_write(micro.geo, "geoexmap/Data_Processed/microplastics.gpkg")
```

## Built Environment

### Walkability
Formula:
w = block group's ranked score for intersection density (D3b)
x = block group's ranked score for proximity to transit stops (D4a)
y = block group's ranked score for employment mix (D2b_E8MixA)
z = block group's ranked score for employment and household mix (D2a_EpHHm)
NWI score = (w/3) + (x/3) + (y/6) + (z/6)
```{r, eval=FALSE}
# let's do our own calculation
# formula:
layers <- st_layers("Data_Raw/WalkabilityIndex/Natl_WI.gdb")

walkability <- st_read("Data_Raw/WalkabilityIndex/Natl_WI.gdb", layer = "NationalWalkabilityIndex")

walkability.wa <- walkability %>% 
  filter(STATEFP == "53")
```

```{r, eval=FALSE}
# get higher level tract fips
walkability.wa$tractFIPS <- substr(walkability.wa$GEOID20, 1, 11)

# group by tract and get mean rankings
walkability.tr <- walkability.wa %>% 
  group_by(tractFIPS) %>% 
  mutate(int_density = mean(D3B_Ranked),
         transit = mean(D4A_Ranked),
         emp_mix = mean(D2B_Ranked),
         emp_hh_mix = mean(D2A_Ranked)) %>% 
  slice(1)

# all good, we have values 1 - 20 now!
walkability.tr$walk_score = (walkability.tr$int_density / 3) +
  (walkability.tr$transit / 3) + (walkability.tr$emp_mix / 6) +
  (walkability.tr$emp_hh_mix / 6)
```

```{r, eval=FALSE}
xwalk <- fread("X:/fast/vopham_t/grp/geoexmap-geoexDC/geoexmap/geoexmap/nhgis_tr2010_tr2020_53.csv")
xwalk$tr2010ge <- as.character(xwalk$tr2010ge)


## join the crosswalk to older geog
wa.xwalk <- merge(xwalk, walkability.tr, by.x = "tr2010ge", by.y = "tractFIPS", all.x = TRUE)

# using block groups for walkability
#caces.wa.xwalk <- merge(xwalk, walkability, by.x = "bg2010ge", by.y = "GEOID", all.x = TRUE)


## multiply older counts/values by crosswalk's interpolation weights, producing estimated counts for all intersections between older and newer geogs
wa.xwalk$walk_score_2020 <- wa.xwalk$wt_pop * wa.xwalk$walk_score

walkability.2020 <- wa.xwalk %>% 
  group_by(tr2020ge) %>% 
  mutate(NatWalkInd2020 = sum(walk_score_2020, na.rm = TRUE)) %>% 
  slice(1) %>% 
  ungroup()

# 0 to 20, recode 15 tracts to have 1

walkability.2020$NatWalkInd2020[walkability.2020$NatWalkInd2020 < 1] <- 1

walkability <- walkability.2020 %>% 
  filter(!is.na(NatWalkInd2020)) %>% 
  data.frame()

walkability.geo <- merge(wa.tracts, walkability, by.x = "GEOID", by.y = "tr2020ge")

walkability <- walkability.geo %>% 
  dplyr::select(c(GEOID, NatWalkInd2020))
```

```{r, eval=FALSE}
st_write(walkability, "geoexmap/Data_Processed/wa_walkability.gpkg", append = FALSE)
```

### Pesticides
```{r, eval=FALSE}
pest <- fread("Data_Raw/V3_Pesticide_Exposure.csv") %>% 
  filter(!`County Name` == "n/a")
```

```{r, eval=FALSE}
pest.geo <- merge(wa.tracts, pest, by.x = "GEOID", 
                  by.y = "2020 Census Tract")

pest.fin <- pest.geo %>% 
  dplyr::select(c(GEOID, pesticide_lbs_mile2, `IBL Rank`))

st_write(pest.fin, "geoexmap/Data_Processed/wa_pest_exposure.gpkg")
```

### Green Space: NDVI
```{r, eval=FALSE}
## NDVI
ndvi <- raster("Data_Raw/Clipped_NDVI_WA_mean.tif")

plot(ndvi)

# convert to spatraster
ndvi_raster <- as(ndvi, "SpatRaster")

plot(ndvi_raster)

# check that projections are the same
st_crs(wa.tracts) == st_crs(ndvi_raster)

# crop
wa_ndvi <- crop(ndvi_raster, extent(wa.tracts))

par(mar=c(1, 1, 1, 1))
plot(wa_ndvi)

# extract mean value of the raster for each census tract
wa.tracts$ndvi <- extract(wa_ndvi, wa.tracts, fun="mean", method = "simple", touches = FALSE,
                           na.rm = TRUE)

wa.tracts.ndvi <- wa.tracts %>% 
  drop_na() %>% 
  mutate(NDVI = ndvi$NDVI) %>% 
  dplyr::select(c(STATEFP, COUNTYFP, TRACTCE, GEOID, NAME, NDVI))

#st_write(wa.tracts, dsn = paste0(getwd(), "/Demo Data/", "wa_tracts_ndvi.gpkg"))

# try RDS for performance
#saveRDS(wa.tracts, "Demo Data/wa_tracts_ndvi_pm25.rds")
```


### Blue Space
```{r, eval=FALSE}
bluespace1 <- rast("Data_Raw/occurrence_130W_50Nv1_4_2021.tif")

plot(bluespace1)

bluespace2 <- rast("Data_Raw/occurrence_120W_50Nv1_4_2021.tif")

plot(bluespace2)

bluespace <- merge(bluespace1, bluespace2)

plot(bluespace)

writeRaster(bluespace, file = "Data_Raw/bluespace.tif")
```

calc proportion area blue space in each tract with 100m buffer
```{r, eval=FALSE}
wa.tract.buffer <- st_buffer(wa.tracts, dist = 100)
```

```{r, eval=FALSE}
plot(wa.tract.buffer$geom)
```

```{r, eval=FALSE}
bluespace.wa <- rast("Data_Raw/bluespace.tif")
```

```{r, eval=FALSE}
plot(bluespace.wa)
```
```{r, eval=FALSE}
# get pixel area
blue.area <- prod(res(bluespace.wa))

# binary raster
bluespace.thresh <- bluespace.wa >= 50

# buffered tracts to terra spatvector
tract.vect <- vect(wa.tract.buffer)
```


```{r, eval=FALSE}
# extract proportion area
# blue.prop <- extract(bluespace.thresh, tract.vect,
#                      fun = mean, na.rm = TRUE, exact = TRUE)
blue.prop <- exact_extract(bluespace.thresh, wa.tract.buffer, 'mean')

wa.tracts$bluespace <- blue.prop
```

```{r, eval=FALSE}
# gut check
library(RColorBrewer)
pal <- colorNumeric("Blues", domain = wa.tracts$bluespace)

leaflet(wa.tracts) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(fillColor = ~pal(bluespace), weight = 0.5,
              color = "black", fillOpacity = 0.7) %>% 
  addLegend(pal = pal, values = ~bluespace, opacity = 0.7, title = NULL, position = "bottomright")
```

```{r, eval=FALSE}
wa.tracts <- wa.tracts %>% 
  dplyr::select(c(GEOID, bluespace))

st_write(wa.tracts, "geoexmap/Data_Processed/wa_bluespace.gpkg")
```

### Light at Night (LAN) (2022)
```{r, eval=FALSE}
lan <- fread("Data_Raw/2024_1031_LAN/VIIRS_DNB/2022/AvgNighttimeRadiance_53.csv")

lan$GEOID <- as.character(lan$GEOID)

# use 2020 geometry
setdiff(lan$GEOID, wa.tracts.2020$GEOID)
setdiff(lan$GEOID, wa.tracts.2019$GEOID)
setdiff(wa.tracts.2019$GEOID, lan$GEOID)
setdiff(wa.tracts.2020$GEOID, lan$GEOID)
setdiff(wa.tracts.2019$GEOID, lan$GEOID)

wa.tracts.lan <- wa.tracts %>% 
  merge(lan, by = "GEOID") %>% 
  dplyr::select(-c(STATEFP.x, STATEFP.y, COUNTYFP.x, COUNTYFP.y, TRACTCE.x,
            TRACTCE.y))

wa.tracts.lan <- wa.tracts.lan %>% 
  dplyr::select(-`.geo`)

saveRDS(wa.tracts.lan, "Demo Data/wa_tracts_pm25_ndvi_lan.rds")

## TODO: process GeoTIFF sent by Seigi
viirs.2022.1 <- raster("Data_Raw/VIIRS_2022_CONUS-0000000000-0000000000.tif")
plot(viirs.2022.1)
viirs.2022.2 <- raster("Data_Raw/VIIRS_2022_CONUS-0000000000-0000032768.tif")
plot(viirs.2022.2)
```


### Noise
```{r, eval=FALSE}
noise <- raster("Data_Raw/CONUS_L50dBA_sumDay_exi.tif")

plot(noise)
```
```{r, eval=FALSE}
# check that projections are the same
st_crs(wa.tracts) == st_crs(noise)

noise <- projectRaster(noise, crs = 4326)

# crop
wa_noise <- crop(noise, extent(wa.tracts))

par(mar=c(1, 1, 1, 1))
plot(wa_noise)

st_crs(wa.tracts) == st_crs(wa_noise)

# convert to spatraster
wa_noise_raster <- as(wa_noise, "SpatRaster")

plot(wa_noise_raster)

# check that projections are the same
st_crs(wa.tracts) == st_crs(wa_noise_raster)


# extract mean value of the raster for each census tract
wa.tracts$noise <- extract(wa_noise_raster, wa.tracts, fun = "mean", method = "simple", touches = FALSE,
                           na.rm = TRUE)

wa.tracts.noise <- wa.tracts %>% 
  drop_na() %>% 
  mutate(noise = noise$CONUS_L50dBA_sumDay_exi) %>% 
  dplyr::select(c(STATEFP, COUNTYFP, TRACTCE, GEOID, NAME, noise))

wa.noise <- wa.tracts.noise %>% 
  dplyr::select(c(GEOID, noise))

st_write(wa.noise, "geoexmap/Data_Processed/wa_noise.gpkg")
```
### Transportation Noise Exposure Model
```{r, eval=FALSE}
# read in
noise.mod <- readRDS("Data_Raw/transportation_noise/tractresultWA.rds") %>% 
  st_drop_geometry()
```

```{r, eval=FALSE}
# 1773 tracts, join
wa.tracts.noisemod <- merge(wa.tracts, noise.mod, by.x = "GEOID", by.y = "GEOID")
```

```{r, eval=FALSE}
# waiting on trang's input
wa.noise <- wa.tracts.noisemod %>% 
  dplyr::select(c(GEOID, noise4050n, noise5060n, noise6070n, noise7080n, noise8090n, noise90n,
           noise4050p, noise5060p, noise6070p, noise7080p, noise8090p, noise90p))
```

```{r, eval=FALSE}
st_write(wa.noise, "geoexmap/Data_Processed/wa_noise_exp.gpkg")
```

### Transit Access
```{r, eval=FALSE}
transit <- st_read("Data_Raw/WSDOT_-_Transit_Stops.shp") %>% 
  dplyr::select(c(stop_name, wheelchair, level6, stop_lat, stop_lon,
           level5, level4, level3, level2,
           level1))

st_write(transit, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "transit", append = FALSE)
```

### Superfund Sites
```{r, eval=FALSE}
superfund_layers <- st_layers("Data_Raw/NPL_Boundaries.gdb")

superfund <- st_read("Data_Raw/NPL_Boundaries.gdb", layer = "SITE_BOUNDARIES_SF")
```
```{r, eval=FALSE}
wa.superfund <- superfund %>% 
  filter(STATE_CODE == "WA")

plot(wa.superfund)
```

```{r, eval=FALSE}
wa.superfund <- wa.superfund %>% 
  st_transform(crs = 4326)
```

```{r, eval=FALSE}
st_write(wa.superfund, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "superfund")
```

### Parks
```{r, eval=FALSE}
us.parks <- st_read("Data_Raw/USA_Parks.gpkg")

wa.state <- st_union(wa.tracts)

wa.parks.filt <- st_contains(wa.state, us.parks, sparse = FALSE) %>% 
  t()

wa.parks <- us.parks %>% 
  filter(wa.parks.filt)

plot(wa.parks$SHAPE)

crs(wa.parks)

st_write(wa.parks, "geoexmap/Data_Processed/wa_parks.gpkg")
```

### Alcohol Retailers
```{r, eval=FALSE}
alc.retail <- fread("Data_Raw/LicenseeListDetails.csv")

alc.retail$Zip <- substr(alc.retail$Zip, start = 1, stop = 5)

colnames(alc.retail)

alc.retail.coords <- alc.retail %>%
  as_tibble() %>% 
  geocode(street = Address, city = City, state = State, postalcode = Zip, method = "arcgis", lat = latitude, long = longitude)

alc.retail.geo <- alc.retail.coords %>% 
  st_as_sf(coords = c("longitude", "latitude")) %>% 
  st_set_crs(4326)

leaflet(alc.retail.geo) %>% 
  setView(lng = -120.74, lat = 47.75, zoom = 7) %>% 
  addMarkers() %>% 
  addTiles() 

st_write(alc.retail.geo, "geoexmap/Data_Processed/wa_alc_retailers.gpkg")
```

### Food Environment

Variables
- POP2010 - population count from 2010 census
- OHU2010 - occupied housing unit count from 2010 census
- lapop1 - low access, pop at 1 mile, number
- lapop1share - low access, pop at 1 mile, share
- lalowi1 - low access, low income pop at 1 mile, number
- lalowi1share - low access, low income pop at 1 mile, share
- lakids1 - low access, children 0-17 at 1 mile, number
- lakids1share - Low access, children age 0-17 at 1 mile, share
- laseniors1 - Low access, seniors age 65+ at 1 mile, number
- laseniors1share - Low access, seniors age 65+ at 1 mile, share
- lawhite1 - Low access, White population at 1 mile, number
- lawhite1share - Low access, White population at 1 mile, share
- lablack1 - Low access, Black or African American population at 1 mile, number
- lablack1share - Low access, Black or African American population at 1 mile, share
- laasian1 - Low access, Asian population at 1 mile, number
- laasian1share - Low access, Asian population at 1 mile, share
- lanhopi1 - Low access, Native Hawaiian and Other Pacific Islander population at 1 mile, number
- lanhopi1share - Low access, Native Hawaiian and Other Pacific Islander population at 1 mile, share
- laaian1 - Low access, American Indian and Alaska Native population at 1 mile, number
- laaian1share - Low access, American Indian and Alaska Native population at 1 mile, share
- laomultir1 - Low access, Other/Multiple race population at 1 mile, number
- laomultir1share - Low access, Other/Multiple race population at 1 mile, share
- lahisp1 - Low access, Hispanic or Latino population at 1 mile, number
- lahisp1share - Low access, Hispanic or Latino population at 1 mile, share
- lahunv1 - Vehicle access, housing units without and low access at 1 mile, number
- lahunv1share - Vehicle access, housing units without and low access at 1 mile, share
- lasnap1 - Low access, housing units receiving SNAP benefits at 1 mile, number
- lasnap1share - Low access, housing units receiving SNAP benefits at 1 mile, share
- TractKids - Total count of children age 0-17 in tract 
- TractSeniors - Total count of seniors age 65+ in tract 
- TractWhite - Total count of White population in tract 
- TractBlack - Total count of Black or African American population in tract 
- TractAsian - Total count of Asian population in tract 
- TractNHOPI - Total count of Native Hawaiian and Other Pacific Islander population in tract 
- TractAIAN - Total count of American Indian and Alaska Native population in tract 
- TractOMultir - Total count of Other/Multiple race population in tract 
- TractHispanic - Total count of Hispanic or Latino population in tract 
- TractHUNV - Total count of housing units without a vehicle in tract 
- TractSNAP - Total count of housing units receiving SNAP benefits in tract 


```{r, eval=FALSE}
healthy.food <- read_excel("Data_Raw/FoodAccessResearchAtlasData2019.xlsx", sheet = "Food Access Research Atlas")

food.access <- healthy.food %>% 
  dplyr::select(c(CensusTract, State, Pop2010, OHU2010, lapop1, lapop1share, lalowi1, lalowi1share, lakids1, lakids1share, laseniors1, 
           laseniors1share, lawhite1, , lawhite1share, lablack1, lablack1share, laasian1, laasian1share, lanhopi1, lanhopi1share,
           laaian1, laaian1share, laomultir1, laomultir1share, lahisp1, lahisp1share, lahunv1, lahunv1share, lasnap1, lasnap1share,
           lasnap1share, TractKids, TractSeniors, TractWhite, TractBlack, TractAsian, TractNHOPI, TractAIAN, TractOMultir, TractHispanic,
           TractHUNV, TractSNAP)) %>% 
  filter(State == "Washington") %>% 
  as.data.frame()

food.access[food.access == "NULL"] <- NA
i <- c(5:41)

food.access[, i] <- sapply(food.access[, i], function(x) as.numeric(x))

# columns 5-29 null in some rows
food.access.geo <- merge(wa.tracts.2019, food.access, by.x = "GEOID", by.y = "CensusTract", all.x = TRUE)

st_write(food.access.geo, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "food_env")
```

### Land Use Land Cover (from Caroline!)
```{r, eval=FALSE}
# read in and take a look
lulc <- st_read("geoexmap/Data_Processed/wa_tracts_NLCD_landcover.gpkg")
```

## Social Environment

### PLACES

### SDOH/Non-Medical Factors

### EJI 
```{r, eval=FALSE}
## EJI
eji <- fread("Data_Raw/EJI_2024_United_States.csv")

eji <- eji %>% 
  filter(STATEFP == 53) %>% 
  mutate(GEOID_2020 = as.character(GEOID_2020)) 
  

wa.tracts.eji <- merge(eji, wa.tracts, by.x = "GEOID_2020", by.y = "GEOID")

st_write(wa.tracts.eji, "geoexmap/Data_Processed/eji.gpkg")
```

### ADI / SVI
ADI Contains some data that are suppressed or unreliable:
- GQ: greater than 33.3% housing units are in group quarters
- PH: population less than 100

Moving forward with SVI
```{r, eval=FALSE}
## SVI (2020 geometry)
svi <- fread("Data_Raw/SVI_WA_2022.csv")
svi$FIPS <- as.character(svi$FIPS)

missing <- svi[is.na(svi$RPL_THEMES)] # length 0

svi <- svi %>% 
  dplyr::select(c(FIPS, RPL_THEMES))

svi.tracts.2022 <- merge(wa.tracts, svi, by.x = "GEOID", by.y = "FIPS") %>% 
  dplyr::select(c(GEOID, RPL_THEMES))

st_write(svi.tracts.2022, "geoexmap/Data_Processed/svi_2022.gpkg", append = FALSE)
```

### Median Household Income
```{r, eval=FALSE}

```

### Racial Residential Segregation (Segregation/Entropy/Dissimilarity Indexs calculation)
```{r, eval=FALSE}
race.eth.bg <- fread("Data_Raw/nhgis0006_csv/nhgis0006_ds258_2020_blck_grp.csv")

race.bg.wa <- race.eth.bg %>% 
  filter(STATEA == 53) %>% 
  dplyr::select(c(GEOCODE, STATEA, COUNTYA, TRACTA, U7J001, U7J002, U7J003, U7J004, U7J005, U7J006, U7J007))

lookup <- c(white = "U7J001", 
            black = "U7J002",
            aian = "U7J003",
            asian = "U7J004",
            nhpi = "U7J005",
            other = "U7J006",
            twoplus = "U7J007")

race.bg <- dplyr::rename(race.bg.wa, all_of(lookup))

race.vars <- race.bg %>% 
  dplyr::select(c(GEOCODE, white, black, aian, asian, nhpi, other, twoplus))
```

```{r, eval=FALSE}
tract_dissim <- race.vars %>% 
  mutate(tractFIPS = substr(GEOCODE, 1, 11)) %>% 
  group_by(tractFIPS) %>% 
  group_modify(~{
    if(nrow(.x) > 1) {
      # calc duncan's between multiple groups
      race_matrix <- as.matrix(.x[, c("white", "black", "aian", "asian", "nhpi", "other", "twoplus")])
      
      dissim_value <- DMulti(race_matrix)
      data.frame(dissim_index = dissim_value)
    } else {
      data.frame(dissim_index = 0)
    }
  }, .keep = TRUE)
```

```{r, eval=FALSE}
seg.geo <- merge(wa.tracts, tract_dissim, by.x = "GEOID", by.y = "tractFIPS")

st_write(seg.geo, "geoexmap/Data_Processed/wa_segregation.gpkg")
```

### Redlining (Sam)
```{r, eval=FALSE}

```

### Location Affordability Index
```{r, eval=FALSE}
lai <- st_read("Data_Raw/Location_Affordability_Index_v3_-8685058346409073024/LOCATION_AFFORDABILITY_INDEX_v3_.shp")
```

### H + T Index
```{r, eval=FALSE}
ht.index <- fread("Data_Raw/htaindex2022_data_tracts_53.csv")
```

```{r, eval=FALSE}
# remove quotes from tract fips column
ht.index$tract <- as.numeric(gsub(".*?([0-9]+).*", "\\1", ht.index$tract))
```

```{r, eval=FALSE}
ht.geo <- merge(wa.tracts, ht.index, by.x = "GEOID", by.y = "tract", all.x = TRUE) %>% 
  dplyr::select(GEOID, ht_ami)
```

```{r, eval=FALSE}
st_write(ht.geo, "geoexmap/Data_Processed/ht_index.gpkg")
```


```{r, eval=FALSE}
# convert to 2020 tracts
lai.coords <- st_coordinates((st_centroid(st_geometry(lai))))
lai.sp <- as(lai, "Spatial")

# add neighbors for disjointed polygons
neighbors <- addnbs(lai.sp)

# neighbors object to matrix
neighbors.q <- spdep::nb2mat(neighbors, zero.policy = TRUE)
```
```{r, eval=FALSE}
# get xwalk
xwalk <- fread("X:/fast/vopham_t/grp/geoexmap-geoexDC/geoexmap/geoexmap/nhgis_tr2010_tr2020_53.csv")
xwalk$tr2010ge <- as.character(xwalk$tr2010ge)

## obtain data of interest for older geog (caces)
# convert to sf
lai <- lai %>% 
  st_as_sf() %>% 
  dplyr::select(c(GEOID, no2_imp, so2_imp, co_imp))

## join the crosswalk to older geog
caces.wa.xwalk <- merge(xwalk, lai, by.x = "tr2010ge", by.y = "GEOID", all.x = TRUE)

# using block groups for walkability
#caces.wa.xwalk <- merge(xwalk, walkability, by.x = "bg2010ge", by.y = "GEOID", all.x = TRUE)


## multiply older counts/values by crosswalk's interpolation weights, producing estimated counts for all intersections between older and newer geogs
caces.wa.xwalk$no2_2020 <- caces.wa.xwalk$parea * caces.wa.xwalk$no2_imp
caces.wa.xwalk$so2_2020 <- caces.wa.xwalk$parea * caces.wa.xwalk$so2_imp
caces.wa.xwalk$co_2020 <- caces.wa.xwalk$parea * caces.wa.xwalk$co_imp

## sum these atom counts for each new census tract (group by the 2020 tract fips)
# there's literally only 1 missing why !
length(unique(xwalk$tr2020ge))
setdiff(wa.tracts$GEOID, xwalk$tr2020ge)

caces.2020.wa <- caces.wa.xwalk %>% 
  group_by(tr2020ge) %>% 
  mutate(no2 = sum(no2_2020),
         so2 = sum(so2_2020),
         co = sum(co_2020)) %>% 
  slice(1) %>% 
  ungroup()

# manually replace: tr2020ge: 53033990100
caces.2020.wa$no2[caces.2020.wa$tr2020ge == 53033990100] <- 0.0
caces.2020.wa$so2[caces.2020.wa$tr2020ge == 53033990100] <- 0.0
caces.2020.wa$co[caces.2020.wa$tr2020ge == 53033990100] <- 0.0

caces.2020.wa <- caces.2020.wa %>% 
  filter(!is.na(no2)) %>% 
  dplyr::select(c(tr2020ge, no2, so2, co))

caces.wa.tracts <- merge(wa.tracts, caces.2020.wa, by.x = "GEOID", by.y = "tr2020ge", all.x = TRUE)
```

### Social Capital
tutorial from: https://rpubs.com/timothyfraser/social_capital_indices
```{r, eval=FALSE}
soc.cap <- read.csv("Data_Raw/index_tracts_V3_04_10_2022.csv")
```

```{r, eval=FALSE}
wa.soc <- soc.cap %>% 
  filter(state == "WA", year == 2020)
```

```{r, eval=FALSE}
# join with tract geo
wa.soc.geo <- merge(wa.tracts, wa.soc, by.x = "GEOID", by.y = "geoid")
```

```{r, eval=FALSE}
# write to file
st_write(wa.soc.geo, "geoexmap/Data_Processed/wa_social_capital.gpkg")
```

### Crimes
```{r, eval=FALSE}
# scratch that, WA county-level NIBRS
wa.crime <- read.csv("Data_Raw/WA_NIBRS_CRIME/Washington_State_Uniform_Crime_Reporting_-_National_Incident_Based_Reporting_System_20250925.csv")
```

```{r, eval=FALSE}
wa.crime.recent <- wa.crime %>% 
  filter(INDEXYEAR == 2023, LOCATION == "COUNTY TOTAL")
```

```{r, eval=FALSE}
# part i and part ii offenses
# part i: homicide, rape, robbery, aggravated assault, burglary, larceny-theft, motor vehicle theft, arson, human trafficking, animal cruelty
# part ii: other assaults (simple), forgery/counterfeiting, fraud, embezzlement, stolen property: buying, receiving, possessing, vandalism, weapons: carrying, possessing, etc., prostitution/commercialized vice, drug abuse, gambling, offenses against family and children, DUI, liquor laws, drunkenness, disorderly conduct, vagrancy, all other offenses, suspicion, curfew/loitering laws
str(wa.crime.recent)
```

```{r, eval=FALSE}
# join with counties
crime.p1 <- wa.crime.recent %>% 
  select(c(COUNTY, POPULATION, PRSNTOTAL, PRSNRATE, SCTYTOTAL, SCTYRATE, PRPRTYTOTAL, PRPRTYRATE, MURDER, MANSLAUGHTER, FORCIBLE_SEX, ROBBERY, ASSAULT, BURGLARY, THEFT, ARSON, HUMAN_TRAFFICKING, KIDNAPPING_ABDUCTION, ANIMAL_CRUELTY))
crime.p2 <- wa.crime.recent %>% 
  select(c(COUNTY, POPULATION, PRSNTOTAL, PRSNRATE, SCTYTOTAL, SCTYRATE, PRPRTYTOTAL, PRPRTYRATE, COUNTERFEITING_FORGERY, DESTRUCTION_OF_PROPERTY, WEAPON_LAW_VIOLATION, PROSTITUTION, NON_FORCIBLE_SEX, BRIBERY, DRUG_VIOLATIONS, GAMBLING_VIOLATIONS, VIOL_OF_NO_CONTACT, EXTORTION_BLACKMAIL))
```

```{r, eval=FALSE}
crime.p1$total_p1 <- rowSums(crime.p1[, c("MURDER", "MANSLAUGHTER", "FORCIBLE_SEX", "ROBBERY", "ASSAULT", "BURGLARY", "THEFT", "ARSON", "HUMAN_TRAFFICKING", "KIDNAPPING_ABDUCTION", "ANIMAL_CRUELTY")])

crime.p2$total_p2 <- rowSums(crime.p2[, c("COUNTERFEITING_FORGERY", "DESTRUCTION_OF_PROPERTY", "WEAPON_LAW_VIOLATION", "PROSTITUTION", "NON_FORCIBLE_SEX", "BRIBERY", "DRUG_VIOLATIONS", "GAMBLING_VIOLATIONS", "VIOL_OF_NO_CONTACT", "EXTORTION_BLACKMAIL")])
```

```{r, eval=FALSE}
# calculate part 1 and part 2 offenses per 1000 people
# number / pop * 1000
crime.p1$p1_rate <- crime.p1$total_p1 / crime.p1$POPULATION * 1000

crime.p2$p2_rate <- crime.p2$total_p2 / crime.p2$POPULATION * 1000
```

```{r, eval=FALSE}
crime.overall <- merge(crime.p1, crime.p2)

crime.overall$COUNTY <- str_to_title(tolower(crime.overall$COUNTY))

crime.geo <- merge(counties, crime.overall, by.x = "NAME", by.y = "COUNTY") %>% 
  select(-c(STATEFP, COUNTYFP, COUNTYNS, AFFGEOID, NAMELSAD, STUSPS, STATE_NAME, LSAD, ALAND, AWATER))
```

```{r, eval=FALSE}
st_write(crime.geo, "geoexmap/Data_Processed/crime.gpkg")
```

### Urbanicity/Rurality (Population Density)
```{r, eval=FALSE}
tot.pop <- st_read("geoexmap/Data_Processed/TotPop_AgeSex.gpkg") %>% 
  dplyr::select(c(1:7))

tot.pop$area <- st_area(tot.pop)

tot.pop$area <- set_units(tot.pop$area, "mi^2")

tot.pop$popdensity <- tot.pop$Total.population / tot.pop$area

pop.density <- tot.pop %>% 
  dplyr::select(c(GEOID, popdensity))

st_write(pop.density, "geoexmap/Data_Processed/wa_popdensity.gpkg", append = FALSE)
```

## Healthcare Access
Minimal processing, point geometry.

### Cancer Programs
```{r, eval=FALSE}
# scrape cancer programs
##html <- read_html("https://www.facs.org/hospital-and-facilities/?searchTerm=&institution=CocHospital&address=wa&sort=relevance&page=3")
cancer.progs <- fread("Data_Raw/cancer_programs_wa.csv")

cancer.prog.coords <- cancer.progs %>%
  as_tibble() %>% 
  geocode(address = Address, method = "arcgis", lat = latitude, long = longitude)

cancer.prog.geo <- cancer.prog.coords %>% 
  st_as_sf(coords = c("longitude", "latitude")) %>% 
  st_set_crs(4326)

leaflet(cancer.prog.geo) %>% 
  setView(lng = -120.74, lat = 47.75, zoom = 7) %>% 
  addMarkers() %>% 
  addTiles() 

st_write(cancer.prog.geo, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg",
         layer = "cancer_progs", append = FALSE)
```

## Data Combining
```{r, eval=FALSE}
# what do we have so far?
setwd("X:/fast/vopham_t/grp/geoexmap-geoexDC/geoexmap/geoexmap")
 
places.social <- st_read("geoexmap/Data_Processed/wa_social.gpkg") %>% 
  st_drop_geometry()
places.prevention <- st_read("geoexmap/Data_Processed/wa_prevention.gpkg") %>% 
  st_drop_geometry()
race.ethnicity <- st_read("geoexmap/Data_Processed/ACS_Race_Ethnicity.gpkg") %>% 
  st_drop_geometry()
pop.age.sex <- st_read("geoexmap/Data_Processed/TotPop_AgeSex.gpkg") %>% 
  st_drop_geometry()
svi <- st_read("geoexmap/Data_Processed/svi_2022.gpkg") %>% 
  st_drop_geometry()
eji <- st_read("geoexmap/Data_Processed/eji.gpkg") %>% 
  st_drop_geometry()
uvi <- st_read("geoexmap/Data_Processed/wa_uvi.gpkg") %>% 
  st_drop_geometry()
places.behaviors <- st_read("geoexmap/Data_Processed/wa_behaviors.gpkg") %>% 
  st_drop_geometry()
places.outcomes <- st_read("geoexmap/Data_Processed/wa_outcomes.gpkg") %>% 
  st_drop_geometry()
radon <- st_read("geoexmap/Data_Processed/wa_radon.gpkg") %>% 
  st_drop_geometry()
walkability <- st_read("geoexmap/Data_Processed/wa_walkability.gpkg") %>% 
  st_drop_geometry()
ozone <- st_read("geoexmap/Data_Processed/wa_ozone_geo.gpkg") %>% 
  st_drop_geometry()
pesticides <- st_read("geoexmap/Data_Processed/wa_pest_exposure.gpkg") %>% 
  st_drop_geometry()
segregation <- st_read("geoexmap/Data_Processed/wa_segregation.gpkg") %>% 
  st_drop_geometry()
noise <- st_read("geoexmap/Data_Processed/wa_noise_exp.gpkg") %>% 
  st_drop_geometry()
sdoh <- st_read("geoexmap/Data_Processed/SDOH_measures.gpkg") %>% 
  st_drop_geometry()
prism <- st_read("geoexmap/Data_Processed/wa_tracts_2020_PRISM.gpkg") %>% 
  st_drop_geometry()
smoke <- st_read("geoexmap/Data_Processed/wildfire_smoke_2023.gpkg") %>% 
  st_drop_geometry()
caces <- st_read("geoexmap/Data_Processed/wa_no2_so2_co.gpkg") %>% 
  st_drop_geometry()
popdensity <- st_read("geoexmap/Data_Processed/wa_popdensity.gpkg") %>% 
  st_drop_geometry()
uvi <- st_read("geoexmap/Data_Processed/wa_uvi.gpkg") %>% 
  st_drop_geometry()
risk <- st_read("geoexmap/Data_Processed/National_Risk_Index.gpkg") %>% 
  st_drop_geometry()
bluespace <- st_read("geoexmap/Data_Processed/wa_bluespace.gpkg") %>% 
  st_drop_geometry()
social.capital <- st_read("geoexmap/Data_Processed/wa_social_capital.gpkg") %>% 
  dplyr::select(c(GEOID, social_capital)) %>% 
  st_drop_geometry()
pfas <- st_read("geoexmap/Data_Processed/wa_dw_pfas.gpkg") %>% 
  st_drop_geometry()
income <- st_read("geoexmap/Data_Processed/med_house_income.gpkg") %>% 
  st_drop_geometry()
ht.index <- st_read("geoexmap/Data_Processed/ht_index.gpkg") %>% 
  st_drop_geometry()
redlining <- st_read("geoexmap/Data_Processed/redlining.gpkg") %>% 
  st_drop_geometry()
lulc <- st_read("geoexmap/Data_Processed/wa_tracts_NLCD_landcover.gpkg") %>% 
  st_drop_geometry()

crime <- st_read("geoexmap/Data_Processed/crime.gpkg") %>% 
  st_transform(crs = 4326)

cancer.progs <- st_read("geoexmap/Data_Processed/wa_cancer_programs.gpkg")
parks <- st_read("geoexmap/Data_Processed/wa_parks.gpkg")
transit <- st_read("Data_Raw/WSDOT_-_Transit_Stops.shp")
alc.retailers <- st_read("geoexmap/Data_Processed/wa_alc_retailers.gpkg")

clinics <- st_read("geoexmap/Data_Processed/clinics.gdb") %>% 
  st_transform(crs = 4326)
ems.stations <- st_read("geoexmap/Data_Processed/ems_stations.gdb") %>% 
  st_transform(crs = 4326)
hospitals <- st_read("geoexmap/Data_Processed/hospitals.gdb") %>% 
  st_transform(crs = 4326)
pharmacies <- st_read("geoexmap/Data_Processed/pharmacies.gdb") %>% 
  st_transform(crs = 4326)
wic.clinics <- st_read("geoexmap/Data_Processed/wic_clinics.gdb") %>% 
  st_transform(crs = 4326)
wic.retailers <- st_read("geoexmap/Data_Processed/WIC_retailers.gdb") %>% 
  st_transform(crs = 4326)

fqhc <- fread("geoexmap/Data_Processed/FQHC.csv") %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269) %>% 
  st_transform(crs = 4326)

microplastics <- st_read("geoexmap/Data_Processed/microplastics.gpkg")

pm25.ndvi.lan <- readRDS("geoexmap/Data_Processed/wa_tracts_pm25_ndvi_lan.rds")

geoexmap.data <- pm25.ndvi.lan %>% 
  left_join(places.social, by = c("GEOID" = "TractFIPS")) %>% 
  left_join(places.prevention, by = c("GEOID" = "TractFIPS")) %>% 
  left_join(race.ethnicity, by = "GEOID") %>% 
  left_join(pop.age.sex, by = "GEOID") %>% 
  left_join(svi, by = "GEOID") %>% 
  left_join(eji, by = c("GEOID" = "GEOID_2020")) %>% 
  left_join(uvi, by = "GEOID") %>% 
  left_join(places.behaviors, by = c("GEOID" = "TractFIPS")) %>% 
  left_join(places.outcomes, by = c("GEOID" = "TractFIPS")) %>% 
  left_join(ozone, by = "GEOID") %>% 
  left_join(radon, by = "GEOID") %>% 
  left_join(walkability, by = "GEOID") %>% 
  left_join(pesticides, by = "GEOID") %>% 
  left_join(segregation, by = "GEOID") %>% 
  left_join(noise, by = "GEOID") %>% 
  left_join(sdoh, by = c("GEOID" = "locationid")) %>% 
  left_join(prism, by = "GEOID") %>% 
  left_join(smoke, by = "GEOID") %>% 
  left_join(caces, by = "GEOID") %>% 
  left_join(popdensity, by = "GEOID") %>% 
  left_join(risk, by = c("GEOID" = "TRACTFIPS")) %>% 
  left_join(bluespace, by = "GEOID") %>% 
  left_join(social.capital, by = "GEOID") %>%  
  left_join(pfas, by = "GEOID") %>% 
  left_join(income, by = "GEOID") %>% 
  left_join(ht.index, by = "GEOID") %>% 
  left_join(redlining, by = "GEOID") %>% 
  left_join(lulc, by = "GEOID") %>% 
  dplyr::select(c(GEOID, pm2.5_t, ndvi_r, avg_rad, FOODSTAMP_CrudePrev, FOODINSECU_CrudePrev, HOUSINSECU_CrudePrev, SHUTUTILITY_CrudePrev,
                  LACKTRPT_CrudePrev, EMOTIONSPT_CrudePrev, ACCESS2_CrudePrev, CHECKUP_CrudePrev, DENTAL_CrudePrev, BPMED_CrudePrev,
                  CHOLSCREEN_CrudePrev, MAMMOUSE_CrudePrev, COLON_SCREEN_CrudePrev, BINGE_CrudePrev, CSMOKING_CrudePrev, LPA_CrudePrev, SLEEP_CrudePrev,
                  ARTHRITIS_CrudePrev, CASTHMA_CrudePrev, BPHIGH_CrudePrev, CANCER_CrudePrev, HIGHCHOL_CrudePrev, COPD_CrudePrev,
                  CHD_CrudePrev, DEPRESSION_CrudePrev, DIABETES_CrudePrev, OBESITY_CrudePrev, TEETHLOST_CrudePrev, STROKE_CrudePrev, Total_population,
                  hispanic_or_latino, pct_hispanic_or_latino, white_non_hispanic, pct_white_non_hispanic, black_non_hispanic, pct_black_non_hispanic,
                  aian_non_hispanic, pct_aian_non_hispanic, asian_non_hispanic, pct_asian_non_hispanic, nhpi_non_hispanic, pct_nhpi_non_hispanic,
                  other_non_hispanic, pct_other_non_hispanic, two_or_more_non_hispanic, pct_two_or_more_non_hispanic, white_hispanic_and_latino,
                  pct_white_hispanic_and_latino, black_hispanic_and_latino, pct_black_hispanic_and_latino, aian_hispanic_and_latino, pct_aian_hispanic_and_latino,
                  asian_hispanic_and_latino, pct_asian_hispanic_and_latino, nhpi_hispanic_and_latino, pct_nhpi_hispanic_and_latino, other_hispanic_and_latino,
                  pct_other_hispanic_and_latino, two_or_more_hispanic_and_latino, pct_two_or_more_hispanic_and_latino, Total.male.population, pct_male,
                  Total.female.population, pct_female,
                  X0_to_4_years_old, pct_0_to_4_years_old, X5_to_9_years_old, pct_5_to_9_years_old, X10_to_14_years_old,
                  pct_10_to_14_years_old, X15_to_19_years_old, pct_15_to_19_years_old, X20_to_24_years_old, 
                  pct_20_to_24_years_old, X25_to_29_years_old, pct_25_to_29_years_old,
                  X30_to_34_years_old, pct_30_to_34_years_old, X35_to_39_years_old, pct_35_to_39_years_old, 
                  X40_to_44_years_old, pct_40_to_44_years_old, X45_to_49_years_old, pct_45_to_49_years_old, 
                  X50_to_54_years_old, pct_50_to_54_years_old, X55_to_59_years_old, pct_55_to_59_years_old, 
                  X60_to_64_years_old, pct_60_to_64_years_old, X65_to_69_years_old, pct_65_to_69_years_old, 
                  X70_to_74_years_old, pct_70_to_74_years_old, X75_to_79_years_old, pct_75_to_79_years_old,
                  X80_to_84_years_old, pct_80_to_84_years_old, X85_years_and_over, pct_85_years_and_over, 
                  RPL_THEMES, RPL_EJI, UVI, radon, NatWalkInd2020, pesticide_lbs_mile2, 
                  dissim_index, noise4050n, noise5060n, noise6070n, noise7080n, noise8090n, noise90n, 
                  noise4050p, noise5060p, noise6070p, noise7080p, noise8090p, noise90p, Unemployment, No.broadband, No.high.school.diploma,
                  Single.parent.households, Crowding, Poverty,
                  Housing.cost.burden, dew, temp_max, temp_min, temp_mean, precip, wildfirePM2.5, no2_imp, so2_imp, co_imp, DS_O3_pred, popdensity, 
                  AVLN_RISKV, CFLD_RISKV, CWAV_RISKV, DRGT_RISKV, ERQK_RISKV, HAIL_RISKV, HWAV_RISKV, HRCN_RISKV, ISTM_RISKV, LNDS_RISKV, LTNG_RISKV, RFLD_RISKV,
                  SWND_RISKV, TRND_RISKV, TSUN_RISKV, VLCN_RISKV, WFIR_RISKV, WNTW_RISKV, bluespace, social_capital, PFAS, Income, ht_ami, HRS2020, pct_Open_Water, pct_Developed_Open, pct_Developed_Low,
                  pct_Developed_Medium, pct_Developed_High, pct_Barren, pct_Evergreen_Forest,
                  pct_Shrub, pct_Grassland, pct_Pasture, pct_Crops, pct_Woody_Wetlands, pct_Herbaceous_Wetlands,
                  pct_Deciduous_Forest, pct_Mixed_Forest, pct_Perennial_Ice))

setnames(geoexmap.data,
         old = names(geoexmap.data),
         new = c("GEOID", "Particulate Matter 2.5", "Green Space", "Nighttime Radiance", "Food Stamps", "Food Insecurity", "Housing Insecurity",
                 "Utility Services Threat", "Lacking Reliable Transportation", "Lack of Social and Emotional Support", "Lack of Health Insurance",
                 "Routine Checkup in the Past Year", "Visited Dentist in Past Year", "Taking Medicine to Control High Blood Pressure",
                 "Cholesterol Screening", "Mammography Use among Women 50 to 74", "Colorectal Cancer Screening among Adults 45 to 75", 
                 "Binge Drinking among Adults", "Cigarette Smoking among Adults", "No Leisure-time Physical Activity among Adults", 
                 "Short Sleep Duration", "Arthritis among Adults", "Asthma among Adults", "High Blood Pressure among Adults", 
                 "Cancer or Melanoma among Adults", "High Cholesterol among Screened Adults", "COPD among Adults", "Coronary Heart Disease among Adults",
                 "Depression among Adults", "Diagnosed Diabetes among Adults", "Obesity among Adults", "All Teeth Lost among Adults 65 and Older", 
                 "Stroke among Adults", "Total Population", "Hispanic or Latino", "Percent Hispanic or Latino", "White NonHispanic",
                 "Percent White NonHispanic", "Black NonHispanic", "Percent Black NonHispanic", "American Indian Alaska Native NonHispanic", 
                 "Percent American Indian Alaska Native NonHispanic", "Asian NonHispanic", "Percent Asian NonHispanic", "Native Hawaiian Pacific Islander NonHispanic", 
                 "Percent Native Hawaiian Pacific Islander NonHispanic", "Other Race NonHispanic", "Percent Other Race NonHispanic", "Two or More Races NonHispanic",
                 "Percent Two or More Races NonHispanic", "White Hispanic or Latino", "Percent White Hispanic or Latino",
                 "Black Hispanic or Latino",  "Percent Black Hispanic or Latino", "American Indian Alaska Native Hispanic or Latino",
                 "Percent American Indian Alaska Native Hispanic or Latino", "Asian Hispanic or Latino", "Percent Asian Hispanic or Latino", 
                 "Native Hawaiian Pacific Islander Hispanic or Latino", "Percent Native Hawaiian Pacific Islander Hispanic or Latino", 
                 "Other Race Hispanic or Latino", "Percent Other Race Hispanic or Latino", 
                "Two or More Races Hispanic or Latino", "Percent Two or More Races Hispanic or Latino", "Total Male Population", "Percent Male", 
                "Total Female Population",  "Percent Female", "Total 0 to 4 years", "Percent 0 to 4 years",
                "Total 5 to 9 years", "Percent 5 to 9 years", "Total 10 to 14 years", "Percent 10 to 14 years", "Total 15 to 19 years", "Percent 15 to 19 years", "Total 20 to 24 years", "Percent 20 to 24 years", "Total 25 to 29 years", "Percent 25 to 29 years", 
                "Total 30 to 34 years","Percent 30 to 34 years", "Total 35 to 39 years", "Percent 35 to 39 years", "Total 40 to 44 years", "Percent 40 to 44 years", "Total 45 to 49 years", "Percent 45 to 49 years", "Total 50 to 54 years", "Percent 50 to 54 years",
                "Total 55 to 59 years", "Percent 55 to 59 years", "Total 60 to 64 years", "Percent 60 to 64 years", 
                "Total 65 to 69 years",  "Percent 65 to 69 years", "Total 70 to 74 years", "Percent 70 to 74 years", "Total 75 to 79 years", "Percent 75 to 79 years",
                "Total 80 to 84 years", "Percent 80 to 84 years", "Total 85 and older", "Percent 85 and older", "Social Vulnerability Index", "Environmental Justice Index", "UV Index", 
                 "Radon", "Walkability", "Pesticide Exposure", "Racial Residential Segregation", "N Noise More than LAeq 45 to 50 db", "N Noise More than LAeq 50 to 60 db", "N Noise More than LAeq 60 to 70 db", "N Noise More than LAeq 70 to 80 db", "N Noise More than LAeq 80 to 90 db", "N Noise More than LAeq 90 db",
                "Pct Noise More than LAeq 45 to 50 db", "Pct Noise More than LAeq 50 to 60 db", "Pct Noise More than LAeq 60 to 70 db",
                "Pct Noise More than LAeq 70 to 80 db", "Pct Noise More than LAeq 80 to 90 db", "Pct Noise More than LAeq 90 db", "Unemployment", 
                 "No broadband internet", "No high school diploma", "Single parent households", "Crowding", "Poverty", "Housing cost burden", 
                 "Dew point", "Maximum temperature", "Minimum temperature", "Average temperature", "Precipitation", "Wildfire smoke", "Nitrogen dioxide",
                 "Sulfur dioxide", "Carbon monoxide", "Ozone", "Population density", "Avalanche Risk Score", "Coastal Flooding Risk Score", "Cold Wave Risk Score",
                "Drought Risk Score", "Earthquake Risk Score", "Hail Risk Score", "Heat Wave Risk Score", "Hurricane Risk Score", "Ice Storm Risk Score", "Landslide Risk Score", "Lightning Risk Score", "Riverine Flooding Risk Score", "Strong Wind Risk Score", "Tornado Risk Score", "Tsunami Risk Score",
                "Volcanic Activity Risk Score", "Wildfire Risk Score", "Winter Weather Risk Score", "bluespace", "social_capital", "PFAS_dw", "Median HH Income", "HT_Index", "Historic Redlining Score",
                "pct_Open_Water", "pct_Developed_Open", "pct_Developed_Low",
                  "pct_Developed_Medium", "pct_Developed_High", "pct_Barren", "pct_Evergreen_Forest",
                  "pct_Shrub", "pct_Grassland", "pct_Pasture", "pct_Crops", "pct_Woody_Wetlands", "pct_Herbaceous_Wetlands",
                  "pct_Deciduous_Forest", "pct_Mixed_Forest", "pct_Perennial_Ice", "geometry"))

st_write(geoexmap.data, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "geoexmap_data", append = FALSE)

st_write(crime, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "county_crime", append = FALSE)

st_write(cancer.progs, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "cancer_progs", append = FALSE)
st_write(parks, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "parks", append = FALSE)
st_write(transit, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "transit", append = FALSE)
st_write(alc.retailers, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "alc_retailers", append = FALSE)

st_write(clinics, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "clinics", append = FALSE)
st_write(ems.stations, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "ems", append = FALSE)
st_write(hospitals, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "hospitals", append = FALSE)
st_write(pharmacies, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "pharmacies", append = FALSE)
st_write(wic.clinics, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "wic_clinics", append = FALSE)
st_write(wic.retailers, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "wic_retailers", append = FALSE)
st_write(fqhc, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "fqhc", append = FALSE)
st_write(microplastics, "geoexmap/Data_Processed/complete/geoexmap_data.gpkg", layer = "microplastics", append = FALSE)
```

