---
title: "Geoexmap Processing"
author: "Malia Cortez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Process and save data to RDS file for use in geoexmap
# MC :)

# Processing:
## Filter raw datasets for just WA census tracts
## Use rasters to extract a mean or other aggregate for each census tract (rather than just using the raw rasters)
## Combine and put all measures into one very wide dataset, where each row is a census tract and contains all corresponding data
## Save to a RDS to conserve memory and use for geoexmap

# quickly loading data
library(data.table)

# tidy and easy data manipulation
library(dplyr)
library(tidyverse)
library(lubridate)

# shiny
library(shiny)
library(bslib)

# LaTeX
#library(tinytex)

# raster/geospatial data manipulation and analysis
library(raster)
library(stars)
library(terra)
library(ncdf4)
library(sf)
library(sp)

# TIGER shapefiles
library(tigris)

# querying Census data
library(tidycensus)

# plotting library for gut checking
library(ggplot2)
library(leaflet)
library(mapdeck)

# set working dir
#setwd("X:/fast/vopham_t/grp/geoexmap-geoexDC/geoexmap")
```

## Current Data
```{r}
# read current data
data <- readRDS("../Data_Processed/wa_tracts_pm25_ndvi_lan.rds")
```

```{r}
# TODO: only take higher level census tracts
data$HL_FIPS <- str_split_i(data$NAME.x, "\\.", 1)

data_HL <- data %>% 
  dplyr::select(-c("NAME.x", `system:index`, "FUNCSTAT", "INTPTLAT", "INTPTLON", "MTFCC", "NAME.y", "NAMELSAD")) %>% 
  group_by(HL_FIPS) %>% 
  summarize(geometry = st_union(geometry), pm2.5 = mean(pm2.5_t), ndvi = mean(ndvi_r), radiance = mean(avg_rad))

plot(data_HL$geometry)
```

```{r}
length(unique(data$NAME.x))
```


```{r}
ex <- c("101.1", "9503.3", "9503", "101.5", "1.002")
str_split_i(ex, "\\.", 1)

```



```{r}
# testing out pivoting for filters in accordion panel
pivoted <- data %>%
  dplyr::select(-c("NAME.x", `system:index`, "FUNCSTAT", "MTFCC", "NAMELSAD", "INTPTLAT", "INTPTLON", "NAME.y")) %>% 
  tidyr::pivot_longer(cols = !c(GEOID, geometry),
               names_to = "Measure",
               values_to = "Value")
```

## Current Map
```{r}
pm25palette <- colorBin(
  palette = "YlOrRd", domain = data$pm2.5_t,
  na.color = "transparent", bins = quantile(data$pm2.5_t)
)

ndvipalette <- colorBin(
  palette = "YlGn", domain = data$ndvi_r,
  na.color="transparent", bins = quantile(data$ndvi_r)
)

lanpalette <- colorBin(
  palette = "OrRd", domain = data$avg_rad,
  na.color="transparent", bins = quantile(data$avg_rad)
)

leaflet(data) %>% 
      setView(lng = -120.74, lat = 47.75, zoom = 7) %>% 
      addTiles() %>% 
      addPolygons(stroke = FALSE, fillOpacity = 0.5, smoothFactor = 0.5, color = ~pm25palette(pm2.5_t), group = "PM 2.5") %>% 
      addPolygons(stroke = FALSE, fillOpacity=0.5, smoothFactor = 0.5, color = ~ndvipalette(ndvi_r), group = "NDVI") %>% 
      #addPolygons(stroke = FALSE, fillOpacity=0.6, smoothFactor=0.5, color = ~pm25palette(wildfirePM2.5), group = "Wildfire PM 2.5") %>%
      addPolygons(stroke = FALSE, fillOpacity=0.5, smoothFactor = 0.5, color = ~lanpalette(avg_rad), group = "LAN") %>% 
      addLayersControl(
        overlayGroups = c("PM 2.5", "NDVI", "LAN"),
        options = layersControlOptions(collapsed = FALSE)
      )
```

## Census Tract Geometry Processing
```{r cars}
# load WA census tracts
#wa.tracts <- tracts(state = "WA", year = 2022, cb = TRUE)
wa.tracts <- tracts(state = "WA", year = 2019, cb = TRUE)
wa.tracts.2020 <- tracts(state = "WA", year = 2020, cb = TRUE)

#st_write(wa.tracts, paste0(getwd(), "/Geo/", "wa_tracts.shp"))

# #wa.tracts <- st_read("../Geo/wa_tracts.shp", promote_to_multi = F) %>% 
#   # switch from NAD83 to WGS84
#   st_transform("+proj=longlat +zone=48 +datum=WGS84") %>% 
#   # erase water from shp
#   erase_water(area_threshold = 0.9) %>% 
#   # ignore linestrings and other polygons
#   st_collection_extract(type = 'POLYGON')

plot(wa.tracts$geometry)

setdiff(wa.tracts.2019$GEOID, smoke$GEOID)
setdiff(smoke$GEOID, wa.tracts.2019$GEOID)

setdiff(wa.tracts$GEOID, smoke$GEOID)
setdiff(smoke$GEOID, wa.tracts$GEOID)

# extract multipolygons

crs(wa.tracts)

wa.tracts$HL_FIPS <- str_split_i(wa.tracts$NAME, "\\.", 1)
length(unique(wa.tracts$GEOID))
length(unique(wa.tracts$HL_FIPS))

length(unique(wa.tracts.2019$GEOID))
length(unique(smoke$GEOID))

wa.tracts <- st_transform(wa.tracts, crs = 4326)

st_write(wa.tracts, "../Geo/wa_tracts_2019.shp", append = FALSE)

# wa.tracts <- wa.tracts %>% 
#   #dplyr::select(-c("NAME.x", `system:index`, "FUNCSTAT", "INTPTLAT", "INTPTLON", "MTFCC", "NAME.y", "NAMELSAD")) %>% 
#   group_by(NAME) %>% 
#   summarize(geometry = st_union(geometry), STATEFP = STATEFP, COUNTYFP = COUNTYFP, )

wa.tracts <- st_read("../Geo/wa_tracts_2019.shp")

```

## PLACES Outcomes, Behaviors, Prevention
```{r}
# ---- OUTCOMES ----
outcomes.2020 <- st_read("../Data_Raw/PLACES__Census_Tract_Data__GIS_Friendly_Format___2020_release_20241119.csv")
```

## Sociodemographics

### ACS
```{r}

```

### ACS - Race
```{r}

```

## Natural Environment
### PM 2.5
```{r natural-env}
#### PM 2.5 ####
pm2.5_demo <- nc_open('Demo Data/V5NA04.02.HybridPM25.NorthAmerica.2022-2022-JUL.nc')

{
  sink('V5NA04.02.HybridPM25.NorthAmerica.2022-2022-JUL.txt')
  print(pm2.5_demo)
  sink()
}

# capture data in lat/long form
lon <- ncvar_get(pm2.5_demo, "lon")
lat <- ncvar_get(pm2.5_demo, "lat", verbose = F)

# capture pm 2.5 data in an array
# measured in microgram/cubic meter
pm2.5_array <- ncvar_get(pm2.5_demo, "GWRPM25")

# 2d array (2022 only)
dim(pm2.5_array)

nc_close(pm2.5_demo)

# save to an array
pm2.5_raster <- raster(t(pm2.5_array), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))

# convert to spatraster
pm2.5_raster <- as(pm2.5_raster, "SpatRaster")

plot(pm2.5_raster)

r <- flip(pm2.5_raster, direction = 'vertical')

plot(r)

# check that projections are the same
st_crs(wa.tracts) == st_crs(pm2.5_raster)

# crop
wa_pm2.5 <- crop(r, extent(wa.tracts))

par(mar=c(1, 1, 1, 1))
plot(wa_pm2.5)

# extract mean value of the raster for each census tract
# 3 missing rows, which have 0.0 population (on water?) and 0 area land (rows 917, 1341, 359)
wa.tracts$PM2.5_t <- extract(wa_pm2.5, wa.tracts, fun="mean", method = "simple", touches = TRUE,
                           na.rm = TRUE)

wa.tracts$PM2.5 <- extract(wa_pm2.5, wa.tracts, fun="mean", method = "simple", touches = FALSE,
                           na.rm = TRUE)

#wa.tracts <- wa.tracts %>% drop_na() %>% 
#  mutate(pm2.5_r = PM2.5$layer) %>% 
#  dplyr::select(c(STATEFP, COUNTYFP, TRACTCE, GEOID, NAME, NAMELSAD, MTFCC, FUNCSTAT, ALAND, AWATER,
#                  INTPTLAT, INTPTLON, pm2.5_r))

wa.tracts <- wa.tracts %>% 
  drop_na() %>% 
  mutate(pm2.5 = PM2.5$layer) %>% 
  dplyr::select(c(STATEFP, COUNTYFP, TRACTCE, GEOID, NAME, pm2.5))

#st_write(wa.tracts, dsn = paste0(getwd(), "/Geo/", "wa_tracts_pm2.5.gpkg"), delete_dsn = TRUE)

#### NITROGEN DIOXIDE NO2 ####

#### OZONE (O3) ####

#### WILDFIRE SMOKE ####
```

### Nitrogen Dioxide (NO2)

### Ozone (O3)
```{r}

```

### Sulfur Dioxide (SO2)
```{r}

```

### Wildfire Smoke / Near Surface Smoke
```{r}
# smoke from Stanford ECHO

smoke.curr <- readRDS("Demo Data/smokePM2pt5_predictions_daily_tract_20060101-20231231.rds")
wa.curr <- smoke.curr %>% 
  filter(startsWith(GEOID, "53"))

smoke2020 <- wa.curr %>% 
  filter(lubridate::year(date) == 2020)

smoke2021 <- wa.curr %>% 
  filter(lubridate::year(date) == 2021)

fwrite(smoke2020, "Demo Data/wa_smoke_2020.csv")
fwrite(smoke2021, "Demo Data/wa_smoke_2021.csv")

smoke20 <- fread("Demo Data/wa_smoke_2020.csv")
smoke21 <- fread("Demo Data/wa_smoke_2021.csv")
smoke22 <- fread("Demo Data/wa_smoke_2022.csv")
smoke23 <- fread("Demo Data/wa_smoke_2023.csv")

smoke20$GEOID <- as.character(smoke20$GEOID)
smoke20$date <- as.Date(smoke20$date)

smoke21$GEOID <- as.character(smoke21$GEOID)
smoke21$date <- as.Date(smoke21$date)

smoke22$GEOID <- as.character(smoke22$GEOID)
smoke22$date <- as.Date(smoke22$date)

smoke23$GEOID <- as.character(smoke23$GEOID)
smoke23$date <- as.Date(smoke23$date)

setdiff(wa.tracts$GEOID, smoke20$GEOID)

length(unique(smoke20$GEOID))
length(unique(smoke21$GEOID))
length(unique(smoke22$GEOID))
length(unique(smoke23$GEOID))

wa.tracts$GEOID <- as.character(wa.tracts$GEOID)

wa.tracts.smoke <- wa.tracts %>% 
  dplyr::select(c("GEOID"))

# processing instructions from: https://www.dropbox.com/scl/fo/91k0aq80vp57qixkm508q/AIRVst386qB2G_N1XhnE4qk/tract?dl=0&preview=README.txt&rlkey=nutebc9pn2vsupr0p9ks4k73u&subfolder_nav_tracking=1
process_smoke <- function(preds) {
  dates <- seq.Date(ymd("20200101"), ymd("20231231"), by = "day")
  
  out <- expand.grid(GEOID = wa.tracts.smoke$GEOID, date = dates) # get full combination of tract days
  out <- left_join(out, preds, by = c("GEOID", "date")) # match smokePM predictions on smoke days to tract days
  
  out <- mutate(out, smokePM_pred = replace_na(smokePM_pred, 0))
  
  return(out)
}

smoke20.full <- process_smoke(smoke20)
smoke21.full <- process_smoke(smoke21)
smoke22.full <- process_smoke(smoke22)
smoke23.full <- process_smoke(smoke23)

# groupby geoid and calculate annual average
smoke_ann_avg <- function(dat) {
  return(dat %>% 
           group_by(GEOID) %>% 
           summarize(wildfirePM2.5 = mean(smokePM_pred)))
}

avg.wildfire2020 <- smoke_ann_avg(smoke20.full)
avg.wildfire2021 <- smoke_ann_avg(smoke21.full)
avg.wildfire2022 <- smoke_ann_avg(smoke22.full)
avg.wildfire2023 <- smoke_ann_avg(smoke23.full)

# join with tract data
wa.tracts.2020 <- merge(wa.tracts, avg.wildfire2020, by = "GEOID")
wa.tracts.2021 <- merge(wa.tracts.2020, avg.wildfire2021, by = "GEOID")
wa.tracts.2022 <- merge(wa.tracts.2021, avg.wildfire2022, by = "GEOID")
wa.tracts.2023 <- merge(wa.tracts.2022, avg.wildfire2023, by = "GEOID")

# rename columns
wa.tracts <- wa.tracts.2023

names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.x"] <- "wildfirePM2.5_2020"
names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.y"] <- "wildfirePM2.5_2021"
names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.x.1"] <- "wildfirePM2.5_2022"
names(wa.tracts)[names(wa.tracts) == "wildfirePM2.5.y.1"] <- "wildfirePM2.5_2023"
```

### UV Radiation
```{r}

```

### AQI
```{r}

```

### Flood Hazard Areas
```{r}

```

### Temperature
```{r}

```

### Precipitation
```{r}

```

### Dew Point
```{r}

```

### Radon 
```{r}

```

### PFAS in Drinking Water
```{r}

```

## Built Environment

### Walkability

### Green Space: NDVI
```{r}
## NDVI
ndvi <- raster("../Data_Raw/Clipped_NDVI_WA_mean.tif")

plot(ndvi)

# convert to spatraster
ndvi_raster <- as(ndvi, "SpatRaster")

plot(ndvi_raster)

# check that projections are the same
st_crs(wa.tracts) == st_crs(ndvi_raster)

# crop
wa_ndvi <- crop(ndvi_raster, extent(wa.tracts))

par(mar=c(1, 1, 1, 1))
plot(wa_ndvi)

# extract mean value of the raster for each census tract
# 3 missing rows, which have 0.0 population (on water?) and 0 area land (rows 917, 1341, 359)
wa.tracts$ndvi <- extract(wa_ndvi, wa.tracts, fun="mean", method = "simple", touches = FALSE,
                           na.rm = TRUE)

wa.tracts <- wa.tracts %>% 
  drop_na() %>% 
  mutate(NDVI = ndvi$NDVI) %>% 
  dplyr::select(c(STATEFP, COUNTYFP, TRACTCE, GEOID, NAME, pm2.5, NDVI))

# m <- leaflet(wa.tracts) %>% 
#   setView(lng = -120.74, lat = 47.75, zoom = 7) %>% 
#   addProviderTiles("OpenStreetMap")
# 
# m <- m %>% addPolygons(
#   stroke = FALSE,
#   fillOpacity = 0.7,
#   smoothFactor = 0.5,
#   color = ~ colorBin("YlGn", ndvi_r)(ndvi_r),
# )
# 
# m

#st_write(wa.tracts, dsn = paste0(getwd(), "/Demo Data/", "wa_tracts_ndvi.gpkg"))

# try RDS for performance
saveRDS(wa.tracts, "Demo Data/wa_tracts_ndvi_pm25.rds")
```


### Blue Space

### Outdoor Light at Night (LAN)
```{r}
lan <- fread("../Data_Raw/2024_1031_LAN/VIIRS_DNB/2022/AvgNighttimeRadiance_53.csv")

wa.tracts.lan <- wa.tracts %>% 
  merge(lan, by = "GEOID") %>% 
  dplyr::select(-c(STATEFP.x, STATEFP.y, COUNTYFP.x, COUNTYFP.y, TRACTCE.x,
            TRACTCE.y))

wa.tracts.lan <- wa.tracts.lan %>% 
  dplyr::select(-`.geo`)

lanpalette <- colorBin(
  palette = "OrRD", domain = wa.tracts.lan$avg_rad,
  na.color="transparent", bins = quantile(wa.tracts.lan$avg_rad)
)

leaflet(wa.tracts.lan) %>% 
  addTiles() %>% 
  addPolygons(stroke = FALSE,
   fillOpacity = 0.7,
   smoothFactor = 0.5,
   color = ~ colorBin("OrRd", avg_rad)(avg_rad)) 

saveRDS(wa.tracts.lan, "Demo Data/wa_tracts_pm25_ndvi_lan.rds")

## TODO: process GeoTIFF sent by Seigi
```


### Noise
```{r}

```

### Transit Access
```{r}

```


## Social Environment

### PLACES

### SDOH/Non-Medical Factors

### EJI (2019 census tract geometry)
```{r}
## EJI
eji <- st_read("../Data_Processed/WA_EJI_2022.shp")

eji <- st_transform(eji, 4326)

leaflet(eji) %>% 
  addTiles() %>% 
  addPolygons(
    stroke = TRUE,
    weight = 1,
    color = "black",
    fillOpacity = 0.7,
    smoothFactor = 0.5,
    fillColor = ~colorBin("YlOrRd", RPL_EJI)(RPL_EJI),
    popup = eji$Location
  )

setdiff(eji$GEOID, wa.tracts.2019$GEOID)
setdiff(eji$GEOID, wa.tracts.2020$GEOID)
setdiff(wa.tracts.2019$GEOID, eji$GEOID)

# try to combine census tracts?
#wa.tracts.new <- wa.tracts.2020 %>% 
#  filter(grepl("[.]", NAME))

# wa.tracts.unmod <- wa.tracts.2020 %>% 
#   filter(!grepl("[.]", NAME))
# 
# wa.tracts.res <- wa.tracts.new %>% 
#   group_by(TRACTCE) %>% 
#   summarize(geometry = st_union(geometry)) %>% 
#   ungroup()
# 
# wa.tracts.new$geometry[match(wa.tracts.res$TRACTCE, wa.tracts.new$TRACTCE)] <- wa.tracts.res$geometry
# 
# wa.tracts.fin <- wa.tracts.new[!duplicated(wa.tracts.new$TRACTCE),]

wa.tracts.eji <- merge()

```

### ADI
```{r}
## ADI
adi.2022 <- fread("../Data_Raw/WA_2022_ADI_Census_Block_Group_v4_0_1.csv")
adi.2020 <- fread("../Data_Raw/WA_2020_ADI_Census_Block_Group_v4_0_1.csv")

adi.2022$StateFips <- substr(adi.2022$FIPS, start = 1, stop = 2)
adi.2022$CountyFips <- substr(adi.2022$FIPS, start = 1, stop = 5)
adi.2022$TractFips <- substr(adi.2022$FIPS, start = 1, stop = 10)

adi.2022$StateFips <- substr(adi.2022$FIPS, start = 1, stop = 2)
adi.2022$CountyFips <- substr(adi.2022$FIPS, start = 1, stop = 5)
adi.2022$TractFips <- substr(adi.2022$FIPS, start = 1, stop = 10)

wa.adi.2022 <- adi.2022 %>% 
  filter(StateFips == "53")

#write.csv("Demo_Data/WA")


#setdiff(eji$GEOID, wa.tracts.2020$GEOID)
```

### Median Household Income
```{r}

```

### Racial Residential Segregation (Segregation/Entropy calculation)
```{r}

```

### Redlining
```{r}

```

### Location Affordability Index
```{r}

```

### Social Capital
```{r}

```

### Crimes
```{r}

```

### Urbanicity/Rurality
```{r}

```

## Healthcare Access
```{r}

```

## Data Combining
```{r}

```

